### 2020-10-03 SingleSite_SingleDisturb.Rmd
Iterative JAGS filter (single site, single disturb) working!

---
title: "Single Site, Single Disturbance"
author: "Michael Dietze"
date: "7/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Initial code for this started from the "Assimilation of disturbance" example in <https://github.com/EcoForecast/EF_BookCode/blob/master/Ch02_Logitic.Rmd> which itself derived from simple simulations and KF developed for a 2013 talk at the Ecological Society of America annual meeting.

### Assimilation of disturbance

"True" run

```{r,echo=1:2, fig.width=7, fig.height=6}
set.seed(1066)
p.dist = 0.05
n0 = 1
nIC = 6
r = .1
K = 6
n.sd = 0.25
NT = 100
time = 1:NT
xlim = c(5,25)
n = rep(nIC,NT)
d = rep(0,NT)
d[10] = 1  ## prescribe a disturbance
for(t in 2:NT){
  n[t] = n[t-1] + r*n[t-1]*(1-n[t-1]/K) + rnorm(1,0,n.sd)
#  d[t] = rbinom(1,1,p.dist)
  if(d[t]) n[t] = abs(rnorm(1,n0,0.5))
}

plot(time,n,type='l',ylim=range(n),xlim=xlim,xlim=xlim,ylab="LAI",lwd=3)
obs.sd = 0.75
n.true=n
n.obs = rnorm(NT,n.true,obs.sd)
points(time,n.obs,pch=18,col=2,cex=1.5)
legend("topleft",legend=c("truth","obs"),lwd=c(2,NA),pch=c(NA,18),col=1:2)
## Simulate less frequent observations
#na.seq = 1:NT; na.seq=na.seq[-seq(1,NT,by=5)]
#n.fia = n.obs;n.fia[na.seq] <- NA
#points(time,n.fia,pch=18,col="blue",cex=2)
```

## "blind" forecast w/o disturbance

```{r}
n0.sd = n.sd
NE = 1000
n0s = rlnorm(NE,log(nIC),n0.sd)
ne0 = matrix(n0s,NE,NT)
for(i in 1:NE){
  for(t in 2:NT){
    ne0[i,t] = abs(ne0[i,t-1] + r*ne0[i,t-1]*(1-ne0[i,t-1]/K) + rnorm(1,0,n.sd))
#    if(rbinom(1,1,p.dist)) ne[i,t] = abs(rnorm(1,n0,0.5))
  }
}
yrng = c(0,8) #range(ne0)
n0.stats = apply(ne0,2,quantile,c(0.025,0.5,0.975))
plot(time,n.obs,type='n',ylim=yrng,xlim=xlim)
polygon(cbind(c(time, rev(time), 1), c(n0.stats[1, ], rev(n0.stats[3, ]),
  n.stats[1, 1])), border = NA, col = 'lightblue')
lines(time,n0.stats[2,],lwd=2,col="blue4")
points(time,n.obs,pch=18,col=2,cex=1.5)
```

## standard EnKF w/o disturbance

```{r,echo=FALSE, fig.width=7, fig.height=6}
## initial ensemble
ens = matrix(NA,NT*2+1,NE)
ens[1,] = rlnorm(NE,log(nIC),n0.sd)
Xae = Pae = 0
Xfe = mean(ens[1,])
Pfe = var(ens[1,])
for(i in 1:NT){

  ##Approximate forecast with Normal
  k = 2*(i-1)+1
  Xfe[i] = mean(ens[k,])
  Pfe[i] = var(ens[k,])

  ## analysis step: combine previous forecast with observed data
  H <- matrix(1,ncol=1)
  Kg <- Pfe[i] * t(H) %*% solve(H%*%Pfe[i]%*%t(H) + obs.sd)
  if(!is.na(n.obs[i])){
    Xae[i] <- Xfe[i] + Kg*(n.obs[i] - Xfe[i]) 
  }else{
    Xae[i] <- Xfe[i]
  }
  Pae[i] <- (1-Kg %*% H)*Pfe[i]

  ## generate new ensemble
  ens[k+1,] = rnorm(NE,Xae[i],sqrt(Pae[i]))

  ## Forecast step: predict to next step from current
  ens[k+2,] = ens[k+1,] + r*ens[k+1,]*(1-ens[k+1,]/K) + rnorm(NE,0,n.sd) 
#  disturbed = as.logical(rbinom(NE,1,p.dist))
#  ens[k+2,disturbed] = abs(rnorm(sum(disturbed),n0,0.5)) ## disturbance

}

## forecast
fx0 = ens[2*(1:NT)-1,]
fx0.stats = apply(fx0,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,fx0.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)  
polygon(cbind(c(time, rev(time), 1), c(n0.stats[1, ], rev(n0.stats[3, ]),      ## Free run, no disturbance
  n0.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(fx0.stats[1, ], rev(fx0.stats[5, ]),    ## Forecast 95%
  fx0.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(fx0.stats[2, ], rev(fx0.stats[4, ]),    ## Forecast IQR
  fx0.stats[2, 1])), border = NA, col = 'green4')
#lines(time[-NT],fx.stats[2,-1],lwd=4,col="green")
lines(time,n0.stats[2,],lwd=2,col="blue4")                                     ## Free run mean
lines(time,n.true,type='l',lwd=2)                                              ## truth
points(time,n.obs,pch=18,col=2)                                                ##obs
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

## analysis
an0 = ens[2*(1:NT),]
an0.stats = apply(an0,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,an0.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)
polygon(cbind(c(time, rev(time), 1), c(n0.stats[1, ], rev(n0.stats[3, ]),
  n0.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(an0.stats[1, ], rev(an0.stats[5, ]),
  an0.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[2, 1])), border = NA, col = 'green4')
lines(time,n0.stats[2,],lwd=2,col="blue4")
lines(time,n.true,type='l',lwd=2)
points(time,n.obs,pch=18,col=2)
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

```

## "blind" forecast w/ disturb

```{r}
n0.sd = n.sd
NE = 1000
n0s = rlnorm(NE,log(nIC),n0.sd)
ne = matrix(n0s,NE,NT)
for(i in 1:NE){
  for(t in 2:NT){
    ne[i,t] = abs(ne[i,t-1] + r*ne[i,t-1]*(1-ne[i,t-1]/K) + rnorm(1,0,n.sd))
    if(rbinom(1,1,p.dist)) ne[i,t] = abs(rnorm(1,n0,0.5))
  }
}
yrng = range(ne)
n.stats = apply(ne,2,quantile,c(0.025,0.5,0.975))
plot(time,n.obs,type='n',xlim=xlim)
polygon(cbind(c(time, rev(time), 1), c(n.stats[1, ], rev(n.stats[3, ]),
  n.stats[1, 1])), border = NA, col = 'lightblue')
lines(time,n.stats[2,],lwd=2,col="blue4")
points(time,n.obs,pch=18,col=2,cex=1.5)
```

## standard EnKF + Disturb

```{r,echo=FALSE, fig.width=7, fig.height=6}
## initial ensemble
ens = matrix(NA,NT*2+1,NE)
ens[1,] = rlnorm(NE,log(nIC),n0.sd)
Xae = Pae = 0
Xfe = mean(ens[1,])
Pfe = var(ens[1,])
for(i in 1:NT){

  ##Approximate forecast with Normal
  k = 2*(i-1)+1
  Xfe[i] = mean(ens[k,])
  Pfe[i] = var(ens[k,])

  ## analysis step: combine previous forecast with observed data
  H <- matrix(1,ncol=1)
  Kg <- Pfe[i] * t(H) %*% solve(H%*%Pfe[i]%*%t(H) + obs.sd)
  if(!is.na(n.obs[i])){
    Xae[i] <- Xfe[i] + Kg*(n.obs[i] - Xfe[i]) 
  }else{
    Xae[i] <- Xfe[i]
  }
  Pae[i] <- (1-Kg %*% H)*Pfe[i]

  ## generate new ensemble
  ens[k+1,] = rnorm(NE,Xae[i],sqrt(Pae[i]))

  ## Forecast step: predict to next step from current
  ens[k+2,] = ens[k+1,] + r*ens[k+1,]*(1-ens[k+1,]/K) + rnorm(NE,0,n.sd) 
  disturbed = as.logical(rbinom(NE,1,p.dist))
  ens[k+2,disturbed] = abs(rnorm(sum(disturbed),n0,0.5)) ## disturbance
#  Xfe[i+1] = m*Xa[i] + beta[2]
#  Pfe[i+1] = q + m*Pa[i]*m

}

## forecast
fx = ens[2*(1:NT)-1,]
fx.stats = apply(fx,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,fx.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)
polygon(cbind(c(time, rev(time), 1), c(n.stats[1, ], rev(n.stats[3, ]),
  n.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(fx.stats[1, ], rev(fx.stats[5, ]),
  fx.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(fx.stats[2, ], rev(fx.stats[4, ]),
  fx.stats[2, 1])), border = NA, col = 'green4')
#lines(time[-NT],fx.stats[2,-1],lwd=4,col="green")
lines(time,n.stats[2,],lwd=2,col="blue4")
lines(time,n.true,type='l',lwd=2)
points(time,n.obs,pch=18,col=2)
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

## analysis
an = ens[2*(1:NT),]
an.stats = apply(an,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,an.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)
polygon(cbind(c(time, rev(time), 1), c(n.stats[1, ], rev(n.stats[3, ]),
  n.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(an.stats[1, ], rev(an.stats[5, ]),
  an.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[2, 1])), border = NA, col = 'green4')
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),  ## Analysis IQR- no disturb
  an0.stats[2, 1])), border = NA, col = 'orange')
lines(time,n.stats[2,],lwd=2,col="blue4")
lines(time,n.true,type='l',lwd=2)
points(time,n.obs,pch=18,col=2)
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

```

## Offline State-Space Model

Start by fitting the full state-space model to the data, then move on to figuring out how to approximate that iteratively

```{r}
library(rjags)

SSSD <- "
model{
  q  ~ dgamma(aq,bq)
  r  ~ dgamma(ar,br)
  R  ~ dgamma(aR,bR)
  K  ~ dgamma(aK,bK)
  p  ~ dbeta(ap,bp)
  for(t in 1:nt){
    Y[t] ~ dnorm(n[t],r)
  }
  n[1] <- N0            ## should add some uncertainty
  for(t in 2:nt){
    muN[t] <- n[t-1] + R*n[t-1]*(1-n[t-1]/K)
    muD[t] ~ dnorm(mu0,pa0)
    D[t] ~ dbern(p)
    mu[t] <- D[t]*muD[t] + (1-D[t])*muN[t]
    n[t] ~ dnorm(mu[t],q)
  }
}"
## Run individual models
data = list(Y=n.obs,nt=NT,
            mu0=1,pa0=1,  ## put a bit more confidence in the low value
            aq=0.1,bq=0.1,
            ar=0.1,br=0.1,
            aR=r,bR=1,
            aK=K,bK=1,
            ap=p.dist*50,bp=(1-p.dist)*50, ## try with an informative prior
            N0 = nIC)  
 mod <- jags.model(file=textConnection(SSSD),
                   data=data,
                  n.adapt=1000,n.chains=3,
                  init=list(R = r, K = K, q = n.sd,p=p.dist))
 jdat <- coda.samples(mod,variable.names=c("R","K","p","r","q"),n.iter=3000) ## burnin
 #plot(jdat)
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=30000,thin=10) ## samples
```

```{r}
## summarize output
CI <- apply(as.matrix(jdat),2,quantile,c(0.025,0.5,0.975))
CIN <- CI[,grep(pattern = "n",colnames(CI))]
col=col2rgb("blue")
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim)
ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
lines(CIN[2,],col="blue",lwd=2)

## Disturbance probability
Dsamp <- as.matrix(jdat)
Dsamp <- Dsamp[,grep(pattern = "^D",colnames(Dsamp))]
muD <- apply(Dsamp,2,mean)

CID <- CI[,grep(pattern = "^D",colnames(CI))]
col=col2rgb("blue")
plot(2:NT,muD,ylim = range(CID),xlim=xlim,type='l')
ecoforecastR::ciEnvelope(2:NT,CID[1,],CID[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

if(FALSE){  ## didn't work as expected because D is 0/1, want to plot p[t]
Dden <- data.frame(val=as.vector(Dsamp),year=rep(1:ncol(Dsamp),each=nrow(Dsamp)))
ggplot2::ggplot(Dden, ggplot2::aes(x = val, y = year,  group = year)) +
  ggridges::geom_density_ridges() +
  ggridges::theme_ridges() + 
  ggplot2::theme(legend.position = "none")
}

## Disturb vs Non
CImuN <- CI[,grep(pattern = "muN",colnames(CI))]
CImuD <- CI[,grep(pattern = "muD",colnames(CI))]
colN=col2rgb("green")
colD=col2rgb("orange")
plot(data$Y,ylim = range(rbind(CImuN,CImuD,data$Y[-1])),xlim=xlim,
     ylab="Disturbance Probability")
ecoforecastR::ciEnvelope(2:NT,CImuD[1,],CImuD[3,],col=rgb(colD[1],colD[2],colD[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(2:NT,CImuN[1,],CImuN[3,],col=rgb(colN[1],colN[2],colN[3],0.3*256,maxColorValue=256))

## comparison to EnKF
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[1, ], rev(an0.stats[5, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[1, ], rev(an.stats[5, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
lines(an0.stats[3,],col="orange",lwd=2)
lines(an.stats[3,],col="#b2df8a",lwd=2)
lines(CIN[2,],col="blue",lwd=2)
points(data$Y,pch=18,col=2)
legend("topright",legend=c("EnKF, no disturb","EnKF, disturb","Multinomial","data"),
       lwd=5,col=c("orange","#b2df8a","blue","red"))

```

## Forward Filter version of SSSD, Known Params

```{r}
FilterSSSD <- "
model{
  N ~ dnorm(muIC,paIC)  ## initial conditions
  ## Forecast
  muN <- N + R*N*(1-N/K)
  muD ~ dnorm(mu0,pa0)
  D ~ dbern(p)
  mu <- D*muD + (1-D)*muN
  n ~ dnorm(mu,q)
  ## Analysis
  Y ~ dnorm(n,r)
}"
## Forecast Cycle
CI.fN = matrix(NA,5,NT)
CI.fD = matrix(NA,5,NT)
Xl = nIC         ## set IC
pl = 1/n.sd^2
Dbar = 0
#mu0 = data$mu0   ## set disturbed state
#pa0 = data$pa0
#aqp = data$aq
#bqp = data$bq
#arp = data$ar
#brp = data$br
for(t in 1:NT){
  ## Forecast + Analysis
  update = list(Y=data$Y[t],
                muIC=Xl,paIC=pl,
                mu0=data$mu0,pa0=data$pa0,
                #aq=aqp[t],bq=bqp[t],ar=arp[t],br=brp[t],
                R=r, K=K,p = p.dist, q = 1/n.sd^2, r=1/obs.sd^2     ## hard coding these to start
                )
  mod <- jags.model(file=textConnection(FilterSSSD),
                   data=update,
                  n.adapt=1000,n.chains=3)#,
#                  init=list(X.ic=Xlf))
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000) 
# plot(jdat)

  ## update parameters  
  dat = as.matrix(jdat)
  Xl  = mean(dat[,"n"])
  pl  = 1/var(dat[,"n"])
  CI.fN[,t] = quantile(dat[,"n"],c(0.025,0.25,0.5,0.75,0.975))
  CI.fD[,t] = quantile(dat[,"D"],c(0.025,0.25,0.5,0.75,0.975))
  Dbar[t] = mean(dat[,"D"])
#  mu0[t+1] = mean(dat[,"muD"])
#  pa0[t+1] = 1/var(dat[,"muD"])

#  mq = mean(dat[,"q"])
#  vq = var(dat[,"q"])
#  aqp[t+1] = mq*mq/vq
#  bqp[t+1] = mq/vq
#  mr = mean(dat[,"r"])
#  vr = var(dat[,"r"])
#  arp[t+1] = mr*mr/vr
#  brp[t+1] = mr/vr
}

```

```{r}
plot(0,0,xlim=xlim,ylim=c(0,1))
ecoforecastR::ciEnvelope(time,CI.fD[1,],CI.fD[5,],col=col.alpha(4,0.5))
ecoforecastR::ciEnvelope(time,CI.fD[2,],CI.fD[4,],col=4)
lines(time,CI.fD[2,])
```

```{r}
## comparison to EnKF
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
#ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(1:NT,CI.fN[2,],CI.fN[4,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

lines(an0.stats[3,],col="orange",lwd=2)
lines(an.stats[3,],col="#b2df8a",lwd=2)
lines(CI.fSSSD[2,],col="blue",lwd=2)
lines(Dbar,lwd=2)
points(data$Y,pch=18,col=2)
legend("topright",legend=c("EnKF, no disturb","EnKF, disturb","Multinomial","data","P(dist)"),
       lwd=5,col=c("orange","#b2df8a","blue","red","black"))
```

## Todo
-   Visualize bimodal forecast and analysis distributions through time for simple example
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Apply to SIPNET SDA