### 2020-11-18 SingleSite_MultiDisturb_MultiPool.Rmd
SSMDMP

---
title: "Single Site, Multiple Disturbance, MultiPool"
author: "Mike Dietze"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BayesianTools)
library(magrittr)
set.seed(42)
## helper functions
source("multi.ens.adj.R")
source("disturbance.R")
```

## Todo

* Multinomial with distinct disturbances
* Multinomial with distinct disturbances that are different PFTs
* Multinomial with initially-overlapping disturbances that then diverge

## SET-UP

```{r}
ne = 200     ## number of ensemble members
NT = 25      ## number of years
nt = 365*NT
time = 1:nt

## DRIVERS
nd   = 31 ## number of driver ensembles
PAR <- matrix(NA,nd,nt)
for(i in 1:nd){
  PAR[i,] <- VSEMcreatePAR(time)  
}

## Parameters
refPars <- VSEMgetDefaults()
refPars$lower[9:11] = 1
np = 50
sigmaP = diag((refPars$upper - refPars$lower)/10)^2 ## assume range is +/- 5 SD
Theta <- as.data.frame(mvtnorm::rmvnorm(np,refPars$best,sigmaP))
colnames(Theta) <- rownames(refPars)
par(mfrow=c(3,4))
for(i in 1:11){
  Theta[Theta[,i] < refPars$lower[i] , i] <- refPars$lower[i]
  Theta[Theta[,i] > refPars$upper[i] , i] <- refPars$upper[i]
  hist(Theta[,i],main=colnames(Theta)[i])
}

## initial conditions
met = sample.int(nd,ne,replace=TRUE)
par = sample.int(np,ne,replace=TRUE)
Na <- matrix(NA,ne,3)
for(i in 1:ne){
  Na[i,] = VSEM(as.numeric(Theta[par[i],]),PAR = PAR[met[i],])[nt,2:4]
}
Na.orig = Na
par(mfrow=c(1,3))
hist(Na[,1],main="Cleaf")
hist(Na[,2],main="Csoil")
hist(Na[,3],main="Cstem")
```

# Multinomial with distinct disturbances

## set up storage, parameters, pseudodata
```{r}
## hard coded 
Q = diag(c(0.1,0.1,0.1)) ## process error
R = diag(c(.25,10,1))      ## observation error
THETA <- as.matrix(Theta[par,]) ## ensemble parameters

## disturbance
mu0 <- matrix(c(0.1,0.1,
                0.1,0.95),2,2)    ## mean leaf & stem biomass after disturbance as % of original
V1  <- diag(c(0.15,0.25))^2       ## var in post disturbance leaf and stem biomass (one for each type)
V2  <- diag(c(0.15,0.01))^2
V0  <- abind::abind(V1,V2,along=3)       ## combine covs
alloc.soil = 0.25        ## fraction of removed C that goes in the soil
p.dist = c(0.9,0.05,0.05)  ## Multinomial transition probability; first is no change

## storage
Dbar = 0
Analysis <- Forecast <- list()
Nmu = NmuK = NfK = Nf = Na*NA

## pseudodata
No = matrix(NA,NT,3)                   ## storage for observed state
Ntrue = matrix(NA,NT,3)                ## storage for latent "true" state
true = 42                              ## reference ensemble member
D.true = rep(0,NT)                     ## true disturbance timeseries
D.true[10] = 1                         ## prescribe a disturbance
refpar = as.numeric(Theta[par[true],]) ## grab the "true" parameters
refpar[9:11] = Na[true,]               ## set initial biomass pools
for(t in 1:NT){
  tsel <- (t-1)*365 + 1:365            ## select 1 yr of met  
  ref.mu = VSEM(refpar,PAR[met[true],tsel])[365,2:4]    ## forward simulation
  ref.f  = mvtnorm::rmvnorm(1,ref.mu,Q)## process error
  if(D.true[t]){
      ref.f = disturbance2(ref.f,type=D.true[t])
  }
  ref.f = tobit(ref.f)
  refpar[9:11] = ref.f                 ## update IC for next iteration
  Ntrue[t,] = ref.f
  No[t,] = tobit(mvtnorm::rmvnorm(1,ref.f,R)) ## observation error
}
par(mfrow=c(3,1))
plot(No[,1])
plot(No[,2])
plot(No[,3])
```

```{r,echo=FALSE,results='hide',message=FALSE}
AnalysisSSMDMP <- "  ## SingleSite MultiDisturbance MultiPool
model{
  ## prior
  for(i in 1:N){
    x[1:3,i]  ~ dmnorm(muf[,i],pf[,,i]) ## forecast priors for different cases
  }
#  rho[] ~ dmulti(p[], 1)          ## disturbance probability
## dmulti throws error, so decomposing by hand
 rho[2] ~ dbern(p[2])
 rho[3] ~ dbin(p[3]/(1-p[2]),1-rho[2])
 rho[1] <- 1 - rho[2] - rho[3]
  ## Analysis 
  X <- rho[1]*x[1:3,1] + rho[2]*x[1:3,2] + rho[3]*x[1:3,3] ## select between cases
#  X <- x%*%rho
  Y ~ dmnorm(X,R)                       ## Observation error
}"

Na <- Na.orig
muf <- matrix(NA,3,3)
pf  <- array(0,c(3,3,3)) ## 3rd variable is class
for(t in 1:NT){

  tsel <- (t-1)*365 + 1:365               ## Select days from driver files

  ## update IC
  THETA[,9:11]  <- Na    ## Multinomial

  ## Forecast 
  for(i in 1:ne){
    fx       = VSEM(THETA[i,], PAR = PAR[met[i],tsel]) ## deterministic simulation
    Nmu[i,]  =  fx[365,2:4]                         ## state variables at time of assim
    Nf[i,]   =  mvtnorm::rmvnorm(1,Nmu[i,],Q)
  }

  ## Disturbance
  d.prior <- rmultinom(ne,1,p.dist)             ## prior on disturbance status
  d.class <- apply(d.prior,2,which.max)
  for(d in 2:3){  ## loop over disturbance classes
    sel.dist <- which(d.class == d)              ## select disturbed ensemble members
    if(length(sel.dist)>3){
      for(j in sel.dist){                      ## run disturbance model
        Nf[j,]  = disturbance2(Nf[j,],type = d-1)
      }
      muf[,d]  = colMeans(Nf[sel.dist,])
      pf[,,d] =  try(solve(var(Nf[sel.dist,])))
    } 
    if(sum(sel.dist) <= 3 || class(pfD) == "try-error") {   ## backup values if the ensemble stats fail
      muf[,d] = colMeans(Nf[d.class==1,])*c(mu0[1,d-1],1,mu0[2,d-1])
      pf[,,d]  = solve(diag(c(0.15,0,0.15)^2))
    }
  }
  ## set undisturbed
  sel.dist = which(d.class == 1)
  muf[,1] = colMeans(Nf[sel.dist,])
  pf[,,1] = solve(var(Nf[sel.dist,]))
  Nf <- tobit(Nf)

  ## priors
  priors <- list(p   = rowMeans(d.prior),         ## disturbance rate
                 muf = muf,    ## forecast means
                 pf  = pf      ## forecast precision
  )
  Forecast[[t]] <- c(list(d.prior=d.prior,Nf=Nf),priors)

  ## Analysis
  update = c(list(Y=No[t,],R=solve(R),N=3),priors)
  mod <- rjags::jags.model(file=textConnection(AnalysisSSMDMP),
                           data=update,
                           n.adapt=1000,n.chains=3)
  jdat <- rjags::coda.samples(mod,n.iter=2000) ## burn-in 
  jdat <- rjags::coda.samples(mod,variable.names=c("X","rho","x"),n.iter=10000)

  ## update parameters & save Analysis posterior
  dat = as.matrix(jdat)
  Xsel <- grep(pattern="^X",x=colnames(dat))

  ## draw new IC
  d.post <- dat[,grep(pattern="rho",x=colnames(dat))]
  d.class.post <- apply(d.post,1,which.max)
  Na <- multi.ens.adj(
          Xf   = Nf,
          cf   = d.class,
          mu.f = t(priors$muf),
          Pf   = abind::abind(    ## find a way to do this with apply(priors$pf,3,solve)
                    solve(priors$pf[,,1]),
                    solve(priors$pf[,,2]),
                    solve(priors$pf[,,3]),
                 along=3),
          Xa = dat[,Xsel],
          ca = d.class.post)
  Ca <- Na[,1] ## updated class assignments
  Na <- tobit(Na[,-1])
#  sel.ic <- sample.int(nrow(dat), ne, replace=TRUE)
#  Na <- dat[sel.ic,nsel]
  Analysis[[t]] <- list(MCMC=dat,
                        Na = Na,
                        Ca = Ca)
}

```

### time series

```{r}
post.dist = sapply(X=Analysis,function(y){
  colMeans(y$MCMC[,grep("rho",colnames(y$MCMC))])
})
plot(post.dist[1,],ylim=c(0,1),type='l')
lines(post.dist[2,],col=2)
lines(post.dist[3,],col=3)
legend("topright",legend=c("undisturbed","harvest","defoliation"),
       lwd=2,col=1:3)

pool.names=c("Cleaf","Csoil","Cstem")
for(i in 1:3){

  CI.Na = sapply(X = Analysis,function(y){
    x = y$MCMC
    sel.n = which(colnames(x) == paste0("X[",i,"]"))
    quantile(x[,sel.n],c(0.025,0.5,0.975))
  })
  #CI.Na = sapply(X = Analysis,function(x){
  #  apply(x,2,quantile,c(0.025,0.975))
  #})

  plot(No[,i],
       #ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[3,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.Na[2,],col="#b2df8a",lwd=2)
  lines(Ntrue[,i],lty=2,lwd=2)
}
```
### Ensemble adj at time of disturbance

```{r,fig.asp=1,fig.width=3}
#par(mfrow=c(1,3))

plot(Analysis[[10]]$Na[,1],Analysis[[10]]$Na[,3],pch=Analysis[[10]]$Ca,col=ecoforecastR::col.alpha("red",alpha=0.5),
     ylim=range(Forecast[[10]]$Nf[,3]),
     xlim=range(Forecast[[10]]$Nf[,1]),
     xlab=pool.names[1],ylab=pool.names[3],
     cex.lab=1.25,asp = 1)
points(Forecast[[10]]$Nf[,1],Forecast[[10]]$Nf[,3],cex=0.5)
points(No[10,1],No[10,3],cex=3,col=ecoforecastR::col.alpha("purple",alpha = 0.75),pch=18)

## ensemble adjustment
sel.a=sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1],Forecast[[10]]$Nf[sel.a,3],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c(1,2,"purple"),
       cex = 1.5)

```