### 2020-11-20 StateTransition.Rmd
initial code for state-and-transition calibration. Not much yet, currently just loading the data

---
title: "State & Transition"
author: "Michael Dietze"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
```

## Objectives

* Use data on land cover, disturbance, and biomass to calibrate a simple state and transition model to drive disturbance probabilities for real-world examples
* set up points and time series for assimilation examples

Metadata: downloaded tile_h03v04 from http://emapr.ceoas.oregonstate.edu/pages/data/viz/index.html on 11/19/2020

## load data

```{r}
# cover
cover = raster("conus_land_cover_ARD_tile_h03v04/conus_land_cover_ARD_tile_h03v04.tif")
extent(cover)
nlayers(cover)
res(cover)
crs(cover)
image(cover)

# disturbance
dis.processed = "dis.h03v04.RData"
if(file.exists(dis.processed)){
  load(dis.processed)
} else {
  ## process disturbance data to match cover's extent and resolution
  dis = raster("Cmonster_disturbance_class_ARD_tile_h03v04/Cmonster_disturbance_class_ARD_tile_h03v04.tif")
  extent(dis)
  nlayers(dis)
  res(dis)
  crs(dis)

  ## match disturbance to cover
  dis <- raster::resample(dis,cover,method="ngb")
  save(dis,file=dis.processed)

  compareRaster(dis, cover, res = TRUE)
  compareRaster(dis,cover,extent = TRUE)
}

```

```{r}
### biomass
#biomass = raster("conus_biomass_ARD_tile_h03v04/conus_biomass_ARD_tile_h03v04.tif")
#extent(biomass)
#res(biomass)
#nlayers(biomass)
#summary(biomass)
```

## estimate cover frequencies, select those of interest
```{r}
f = freq(cover)
## most frequent class is 42 = evergreen forest

nbands(cover)
b = bandnr(cover)

cvr = getValues(cover)
cvr = as.matrix(cover[[1]])

```

## Estimate disturbance probability for each cover class
```{r}
fd = freq(dis)

#### disturbance aggregations ####

## 
dis[dis==60] <-

```

## Determine transitions (cover / pft)

## Reduce dimensionality, save

## select sites of interest, save