### 2020-12-23 OR_original_model.Rmd
Code and data extractions to start doing test of the model against real data (landtrendr)

---
title: "OR Multi Disturbance"
author: "Michael Dietze"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("disturbance.R")
source("multi.ens.adj.R")
```

## set up
```{r}
## Load disturbance rate, state&transition, disturbed biomass stats
load("STprior.RData")
colnames(dextract) <- paste0("t",dtime)
dextract[is.na(dextract)] <- 0
d.prior["UN"] = d.prior["UN"] + d.prior[6]                         ## change at runtime (removes <NA>)
d.prior = d.prior[-6]

## Load Calibration
load("fit.RData")

## Load extracted site timeseries
load("extract.h03v04.RData")
```

## drivers
```{r}
latlon = points@coords
metFile = "daymet.RData"
if(file.exists(metFile)){
  load(metFile)
} else {
  dm = list()
  for(i in 1:nrow(latlon)){
    dm[[i]] = daymetr::download_daymet(site="h03v04",latlon[i,2],latlon[i,1],start=1985,end=2017,internal=TRUE,simplify=TRUE)
  }
  save(dm,file=metFile)
}

## extract Shortwave and convert to daily PAR
SW2PAR = 0.45
srad = subset(dm[[1]],measurement == "srad..W.m.2.")   
metTime <- as.POSIXct(paste(srad$year,srad$yday,sep="-"),format="%Y-%j")
rad <- matrix(NA,length(dm),length(metTime))
for(i in seq_along(dm)){
  srad = subset(dm[[i]],measurement == "srad..W.m.2.")   
  rad[i,] = srad$value*1e-6*86400*SW2PAR  # -> MJ /m2 /day PAR
}
plot(metTime,rad[1,],type='l')
## plot(date,PAR,col=2) ## calibration data
```

## functions
```{r,echo=FALSE,results='hide',message=FALS}
##' Single Site, Single Disturbance
##' @param Y observed data
##' @param R observation error
##' @param Q process error
##' @param Xic state initial condition matrix
##' @param THETA parameter matrix
##' @param dist  disturbance list: p.dist, mu0[3], V0[3,3], alloc soil 
SSSD.model <- function(Y,R,Q,PAR,Xic,THETA,dist){

  ## init
  NT = length(PAR)/365
  Nf = Nmu = Na = Xic
  ne = nrow(Xic)
  if(!is.matrix(THETA)) THETA = matrix(THETA,nrow=ne,ncol=length(THETA),byrow = TRUE)
  Forecast = Analysis = list()

  AnalysisSSSDMP <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      muN ~ dmnorm(mufN,pfN)                ## forecast prior on Not disturbed
      muD ~ dmnorm(mufD,pfD)                ## forecast prior on Disturbed
      D ~ dbern(p)                          ## disturbance probability
      ## Analysis 
      n <- D*muD + (1-D)*muN                ## select between disturbed and undisturbed
      Y[H] ~ dmnorm(n[H],R[H,H])                       ## Observation error
  }"

  for(t in 1:2){

    ## update IC
    THETA[,9:11]  <- Na    ## Multinomial

    ## Forecast 
    for(i in 1:ne){
      tsel    = (t-1)*365 + 1:365             ## Select days from driver files
      Nmu[i,] = VSEM(THETA[i,], PAR = PAR[tsel])[365,2:4] ## deterministic simulation
      Nf[i,]  = mvtnorm::rmvnorm(1,Nmu[i,],Q) ## process error
    }

    ## Disturbance
    d.forecast <- rep(0,ne); d.forecast[1:ceiling(ne*0.1)] <- 1 #rbinom(ne,1,dist$p.dist)             ## prior on disturbance status
    sel <- !as.logical(d.forecast)                ## select undisturbed ensemble members
    if(sum(!sel)>2){
      sel.dist = which(!sel)
      for(j in sel.dist){                      ## run disturbance model
        Nf[j,]  = disturbance.p(Nf[j,],mu0=dist$mu0,V0 = dist$V0,alloc.soil = dist$alloc.soil)
      }
      mufD  = colMeans(Nf[!sel,])
      pfD = solve(var(Nf[!sel,]))#solve(diag(c(0.15,0.5,0.25)^2))
    } 
    if(sum(!sel) <= 2 || class(pfD) == "try-error") {   ## backup values if the ensemble stats fail
      mufD = c(dist$mu0[1], mean(Nf[,2]),  dist$mu0[2])            
      pfD  = solve(diag(c(0.15,0.5,0.25)^2))
    }
    Nf <- tobit(Nf)

    ## priors
    priors <- list(p    = dist$p.dist,         ## disturbance rate
                   mufN = colMeans(Nf[sel,]),    ## Not disturbed forecast means
                   mufD = mufD,   ## Disturbed forecast means
                   pfN  = solve(var(Nf[sel,])),  ## Not disturbed forecast precision
                   pfD  = pfD  ## Disturbed forecast precision
    )
    Forecast[[t]] <- c(list(d.forecast=d.forecast,Nf=Nf),priors)

    ## Analysis
    update = c(list(Y=Y[t,],R=solve(R)),priors,H = which(!is.na(Y[t,])))
    if(length(update$H) > 0){ ## data observed, run filter
      mod <- rjags::jags.model(file=textConnection(AnalysisSSSDMP),
                               data=update,
                               n.adapt=1000,n.chains=3,quiet=TRUE)
      jdat <- rjags::coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000)

      ## update parameters & save Analysis posterior
      dat = as.matrix(jdat)
      nsel <- grep(pattern="^n",x=colnames(dat))
      Xa = dat[,nsel]
      ca = dat[,"D"]
      run.adjustment = TRUE
      if(FALSE){
        d3 <-density(Xa[,3]) 
        plot(d3)
        lines(d3$x,dnorm(d3$x,update$Y[3],1/sqrt(R[3,3])),col=2)
        lines(d3$x,dnorm(d3$x,priors$mufN[3],1/sqrt(priors$pfN[3,3])),col="blue")
        p.kf = R[3,3]+priors$pfN[3,3]
        mu.kf = priors$mufN[3]*priors$pfN[3,3]/p.kf+
          update$Y[3]*R[3,3]/p.kf
        lines(d3$x,dnorm(d3$x,mu.kf,1/sqrt(p.kf)),col=3,lty=2)

        jtest <- "model{
         Y ~ dmnorm(n[H],R[H,H])
        }"
        jtest <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      n[3] ~ dnorm(mufN[3],pfN[3,3])  
      Y[H] ~ dnorm(n[H],R[H,H])
      dummy <- mufD[1] + pfD[1,1] + p
  }"
      jtest <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      n[3] ~ dnorm(mufN[3],pfN[3,3])  
      Y[H] ~ dnorm(n[3],R[H,H])
  }"
        update2 = list(mufN = update$mufN,
                       pfN = update$pfN,
                       Y=update$Y,R=R,H=3)
        mtest <- rjags::jags.model(file=textConnection(jtest),
                                   data = update,
#                               data=list(n = c(NA,NA,mu.kf),
#                                         R=matrix(p.kf,3,3),H=3),
                               n.adapt=1000,n.chains=3,quiet=TRUE)
        mdat <- rjags::coda.samples(mtest,variable.names=c("n"),n.iter=10000) 
      summary(mdat)  
      }
    } else {
      Xa = Nf
      ca = 1
      run.adjustment = FALSE
    }

      ## draw new IC
    if(run.adjustment){
    Na <- multi.ens.adj(
      Xf   = Nf,
      cf   = d.forecast,
      mu.f = rbind(priors$mufN,priors$mufD),
      Pf   = abind::abind(solve(priors$pfN),
                            solve(priors$pfD),
                            along=3),
      Xa = Xa,
      ca = ca)
    } else {
      Na = cbind(rep(0,nrow(Nf)),Nf)
    }
    Ca <- Na[,1] ## updated classf assignments
    Na <- tobit(Na[,-1])
    #  sel.ic <- sample.int(nrow(dat), ne, replace=TRUE)
    #  Na <- dat[sel.ic,nsel]
    Analysis[[t]] <- list(Na = Na,
                          Ca = Ca,
                          CI = apply(dat,2,quantile,c(0.025,0.25,0.5,0.75,0.975))
    )
  }
  return(list(Forecast=Forecast,Analysis=Analysis))
}
```

## initial conditions & observations
```{r}
## BIOMASS
B2C = 0.5 ##biomass to carbon ratio
Yw <- bextract*B2C*1000*1e-4 ## convert Mg/ha  to  kg C 
Rw <- (84*B2C*1000*1e-4)^2 ## Observation Error: Kennedy et al 2018 Environ. Res. Lett. 13 025004; Fig 2 RMSE
## units: Mg/ha -> kg/m2

## grab the distribution of AGB for the first time point from all pixels
AGBprior <- Yw[,1]
ne = length(AGBprior) ## number of ensemble members

## spin up
PARlong <- as.vector(t(rad))
spin = VSEM(theta,PARlong)
thin = seq(nrow(spin)/2,nrow(spin),length=ne)
spinIC = spin[thin,2:4]
#for(i in 2:4){
#  hist(spin[thin,i],type='l')
#}

## disturbance recovery trajectory
nNBG = 200 ## years
NBG <- array(NA,c(ne,nNBG,3))
for(i in 1:ne){
  param = theta
  param[9:11] <- spinIC[i,]*c(0.01,1,0.01) ## set leaf and wood to 1%
  raw = VSEM(param,PARlong[1:(nNBG*365.25)])
  NBG[i,,] = raw[round(seq(365,nNBG*365.25,length=nNBG)),2:4]  ## grab pools at the end of each year
}

## for each AGBprior, grab the matching leaf and soil
ICprior <- matrix(NA,ne,3)
age = rep(NA,ne)
for(i in 1:ne){
  age[i] <- which.min((AGBprior[i]-NBG[i,,3])^2)  
  ICprior[i,] <- c(NBG[i,age[i],1:2],AGBprior[i])
}
hist(age)
colnames(ICprior) <- c("Cv","Cs","CR") 
pairs(ICprior)

```

## analyze sites - just one disturbance
```{r}
i = 181  # 177:263 = #fire
NT = floor(ncol(rad)/365)
Yi = matrix(NA,NT,3)  ## observed data
Yi[1:ncol(Yw),3] = Yw[i,]  ## only Cw observed
R = diag(rep(Rw/1000,3))    ## observation error   ### div 1000 is test
Q = diag(c(Ql,Qs,Qw)^2)*10  ## process error
dist <- list(              ## disturbance parameters
  p.dist = d.prior[4],     ## disturbance probability
  mu0 = c(0.2,PBBAR[4,4]),   ## % retention [leaf,agb], first parameter made up, should include soil
  V0 =  diag(c(0.15,PBVAR[4,4])^2), #variance in retention
  alloc.soil = 0.05        ## made up  
)
test = SSSD.model(Y = Yi,R=R,Q=Q,
                  PAR=rad[i,],Xic=ICprior,THETA=theta[1:11],
                  dist = dist)
```

## visualization
```{r}
i = 181
pool.names=c("Cleaf","Csoil","Cstem")
for(i in 1:3){

    CI.Na = sapply(X = test$Analysis,function(x){
      y = x$CI
      sel.n = which(colnames(y) == paste0("n[",i,"]"))
      y[,sel.n]
    })

    CI.D = sapply(X = test$Analysis,function(x){
      y = x$CI
      sel.n = which(colnames(y) == paste0("D"))
      y[,sel.n]
    })

    CI.Nf = sapply(X = test$Forecast, function(x){
      mu = x$mufN[i]
      sd = 1/sqrt(x$pfN[i,i])
      return(mu + c(-1.96,0,1.96)*sd)
    })

  plot(Yi[,i],
       ylim = range(rbind(CI.Na,Yi[,i]),na.rm=TRUE),
       type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.Nf),CI.Nf[1,],CI.Nf[3,],col=ecoforecastR::col.alpha("blue",0.5))
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[5,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.Na[3,],col="#b2df8a",lwd=2)
  points(Yi[,i])
  lines(CI.D[3,])

}

```

 TODO:
 * why is Analysis not contracting around the data?
    * reducing obs error by /10 didn't solve
 * why are the disturbance forecast pools crazy?
 * assess disturbance posterior

## Computer experiments
* Run under defaults
* Run under a grid of inc process error and decrease obs error and increase disturbance probability

## postprocessing

false positives

false negatives

classification success

## Next steps:
* Run with multiple disturbance options and PFT switching