### 2020-12-23 calibrate.Rmd

---
title: "Calibration"
author: "Michael Dietze"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BayesianTools)
citation("BayesianTools")
set.seed(42)
```

## Data stream 1: Eddy covariance flux & met
```{r}
AF = read.csv("Metolius.csv",header=TRUE)

AF.biomass = raster::extract(bioS,AF[,c("Lat.","Long.")])

LAI.BADM = c(1.87, 1.96, 2.35, 2.37, 2.7, 2.8, 2.87, 3.01)

SoilCcon = c(0.37192982456,0.18974358974,0.10518518519)  ##gC/kgSoil
SoilBD   = c(1.14,1.17,1.35)/1000 ## kgSoil/cm3  
SoilDepth= c(20,30,50)  ## cm
SoilC = sum(SoilCcon*SoilBD*SoilDepth*1e4) ## gC/kgSoil * kgSoil/cm3 * cm3/m2  ->kgC/m2

flux = read.csv("Ameriflux/AMF_US-Me2_BASE-BADM_14-5/AMF_US-Me2_BASE_HH_14-5.csv",header = TRUE,skip=2,na.strings = c(-9999,-6999))
time = as.POSIXlt(as.character(flux$TIMESTAMP_START),format="%Y%m%d%H%M")

## PAR data
ein = 2.17e-5 #J mol-1 from ED2 wiki
ein = 0.327 * 1e6 #J/umol * 1e6 umol/mol #units from Apogee 
date.i = as.Date(time)
date = as.Date(names(tapply(date.i,date.i,mean)))
PAR = tapply(flux$PPFD_IN,date.i,mean)*ein*86400*1e-12 ## µmolPhoton m-2 s-1 -> MJ /m2 /day

## Gapfill
doy <- tapply(lubridate::yday(time),date.i,mean)
PARclim <- tapply(PAR,doy,mean,na.rm=TRUE)
PAR[is.na(PAR)] <- PARclim[doy[is.na(PAR)]]

plot(date,PAR)

## NEE
NEE = tapply(flux$FC,date.i,mean,na.rm=TRUE)
NEEp = tapply(flux$NEE_PI,date.i,mean,na.rm=TRUE)
cnt = tapply(is.na(flux$FC),date.i,sum)
cntp = tapply(is.na(flux$NEE_PI),date.i,sum)
NEE2 = NEE
NEE2[cnt<1] = NA
#plot(time,flux$FC)
plot(date,NEE)
points(date,NEE2,col=2)

plot(doy,NEE)
points(doy,NEE2,col=2)

## aggregate flux to annual and change units to kgC/m2/year
years <- lubridate::year(date)
aNEE = tapply(NEE,years,mean,na.rm=TRUE)*1.2e-8 * 86400 *365 ## umol/m2/sec * kg/umol * sec/day
year = as.numeric(names(aNEE))
plot(year,aNEE)

## leaf and wood biomass
Cw = Cl = rep(NA,19)
Cl[c(1,2,3,4,9,10,5,6,7,8)] = c(452.80554869,460.10972745,469.24831463,477.98289883,525.91518937,535.12650483, 485.76130987,496.26140507,506.1776831,514.03435729)/1000
Cw[c(1,2,3,4,10,5,6,7,8,9)] = c(4435.1119994,4793.8053523,4968.8831826,5140.6894352,6293.0243086,5305.4365914,5504.2730671,5699.6935828,5871.0926993,6100.944853)/1000

```

## Data stream 2: LAI

## Data stream 3: AGB

## Data stream 4: spatial meteorology

## Define Likelihood, priors
```{r}
parSel = c(3, 4, 6, 7, 8, 12:14)

## Priors
refPars <- VSEMgetDefaults()
refPars[12,] <- c(.1, 0.01, 4)  ## sigma NEE
refPars[13,] <- c(.2, 0.01, 4)  ## sigma Cl
refPars[14,] <- c(.5, 0.01, 4)  ## sigma Cw
rownames(refPars)[12:14] <- c("sd_NEE","sd_leaf","sd_wood")
refPars["LUE","best"] <- 0.00022 ## tuning
refPars["GAMMA",] <- c(0.55,0.2,0.7) ## tuning
refPars["tauS",] <- c(5e+04,4000,1e5) 
refPars["LAR","best"] <- 1000/127/0.5   ## Ameriflux BADM says LMA is 127 g m-2  -> m2/kgC
refPars["tauV","best"] <- 4.7*365.25 ## Wikipedia says var. ponderosa has a longevity of 4.7+/-0.14 years
refPars["tauR",] <- c(100,5,400)*356 ## sylvics says 300-year lifespan, but there will be branchfall and fine litter
refPars["Av","best"] <- 113.81026786/(113.8 + 206.2)*0.95 ## Ameriflux litter prod/(litter+wood prod) * tuning fudge factor
refPars["Cv","best"] <- Cl[1]*1.5-0.5*Cl[2] #0.4528 ## Ameriflux gC/m2 -> kgC/m2 ## LAI.BADM[1]/refPars["LAR","best"]
refPars["Cr","best"] <- Cw[1]*1.5-0.5*Cw[2]#4.4351119994 ## Ameriflux
refPars["Cs","best"] <- SoilC/10  ## BADM value seems too high
prior <- createUniformPrior(lower = refPars$lower[parSel], 
                            upper = refPars$upper[parSel], best = refPars$best[parSel])

## Likelihood

# here is the likelihood 
likelihood <- function(par, sum = TRUE){
  # set parameters that are not calibrated on default values 
  x = refPars$best
  x[parSel] = par
  pred <- VSEM(x[1:11], PAR) # replace here VSEM with your model
  NEEm = tapply(pred[,"NEE"],years,sum)
  Cl.m  = tapply(pred[,"Cv"],years,mean)
  Cw.m  = tapply(pred[,"CR"],years,mean)
  llValues <- c(0.25*dnorm(x = aNEE, mean = NEEm, sd = x[12], log = TRUE),  ## downweight because less important
              dnorm(Cl,Cl.m,x[13],log=TRUE),
              4*dnorm(Cw,Cw.m,x[14],log=TRUE)) ## upweight (destroys prob interpretation, but var most important)
  llValues[is.nan(llValues)]<- -Inf
  return(sum(llValues,na.rm = TRUE))
}

# optional, you can also directly provide lower, upper in the createBayesianSetup, see help
bayesianSetup <- createBayesianSetup(likelihood, prior, 
                                     names = rownames(refPars)[parSel])#,parallel = TRUE)#,parallelOptions = list(variables=list(refPars=,"parSel","PAR","years","aNEE","Cl","Cw")))
```

## Precalibration runs
```{r}
pred = VSEM(refPars[1:11,"best"],PAR = PAR)
NEEm = tapply(pred[,"NEE"],years,sum)

plot(NEEm,aNEE)
abline(0,1,col=2,lty=2,lwd=3)

#plot(doy,NEE)
#points(doy,pred[,"NEE"],pch="+",col=2)

plot(year,tapply(pred[,"Cv"],years,mean))
points(year,Cl,col=2,pch="+")

plot(year,tapply(pred[,"CR"],years,mean))
points(year,Cw,col=2,pch="+")

plot(pred[,"Cs"])

```

## Fit via MLE
```{r}
negLL <- function(par){-1*likelihood(par)}
mle = optim(refPars[parSel,"best"],negLL)
#mle = optim(refPars[parSel,"best"],negLL,method="L-BFGS-B",lower=refPars[parSel,"lower"],upper=refPars[parSel,"upper"])
cbind(refPars[parSel,"best"],mle$par)

likelihood(refPars[parSel,"best"])
likelihood(mle$par)

```

## Fit via Bayesian Tools
```{r}
settings <- list(iterations = 25000, nrChains = 2)#,parallel=TRUE)
out <- runMCMC(bayesianSetup = bayesianSetup, sampler = "DEzs", settings = settings)
foo = summary(out)
```

```{r}
plot(out)
```

##postcalibrartion
```{r}
ref = refPars[,"best"]
#fit = BayesianTools::MAP(out)$parametersMAP
fit = mle$par
ref[12:14] = fit[6:8]
likelihood(ref[parSel])
likelihood(fit)

theta = refPars[,"best"]
cbind(theta[parSel],fit)
theta[parSel] = fit
names(theta) <- rownames(refPars)
post = VSEM(theta[1:11],PAR = PAR)
NEEm = tapply(post[,"NEE"],years,sum)

plot(NEEm,aNEE)
abline(0,1,col=2,lty=2,lwd=3)

#plot(doy,NEE)
#points(doy,pred[,"NEE"],pch="+",col=2)

Cl.pred = tapply(pred[,"Cv"],years,mean)
Cl.post = tapply(post[,"Cv"],years,mean)
plot(year,Cl.pred,ylim=range(c(Cl,Cl.post,Cl.pred),na.rm=TRUE))
lines(year,Cl.post,col=3)
points(year,Cl,col=2,pch="+")

Cw.post = tapply(post[,"CR"],years,mean)
plot(year,tapply(pred[,"CR"],years,mean))
lines(year,Cw.post,col=3)
points(year,Cw,col=2,pch="+")

Cs.post = tapply(post[,"Cs"],years,mean)
plot(pred[,"Cs"])
lines(Cs.post)
abline(h=SoilC/10)

## Rough approximation of process error: RMSE on delta
Ql <- sqrt(mean((diff(Cl)-diff(Cl.post))^2,na.rm=TRUE))
Qw <- sqrt(mean((diff(Cw[-1])-diff(Cw.post[-1]))^2,na.rm=TRUE)) ## first value bit of an outlier
Qs <- sqrt(mean(diff(Cs.post))) ## assume 100% error (i.e. true delta Soil C = 1)

## Cv of Q
Ql/mean(diff(Cl),na.rm=TRUE)
Qw/mean(diff(Cw),na.rm=TRUE)

save(theta,Ql,Qw,Qs,file="fit.RData")
```

## Ameriflux Download Manifest
```
#downloadDate:20201206113353
#numberOfSitesDownloaded:6
#emailForSitePIs:bev.law@oregonstate.edu
#dataFiles
SITE_ID,filename,start_year,end_year,version,filetype,timestamp
US-Me1,AMF_US-Me1_BASE_HH_3-5.csv,2004,2005,3-5,FLUX-MET,201902011245
US-Me2,AMF_US-Me2_BASE_HH_14-5.csv,2002,2020,14-5,FLUX-MET,202011171249
US-Me3,AMF_US-Me3_BASE_HH_4-5.csv,2004,2009,4-5,FLUX-MET,201811191702
US-Me4,AMF_US-Me4_BASE_HH_4-1.csv,1996,2000,4-1,FLUX-MET,201602060503
US-Me5,AMF_US-Me5_BASE_HH_2-1.csv,1999,2002,2-1,FLUX-MET,201605150147
US-Me6,AMF_US-Me6_BASE_HH_12-5.csv,2010,2020,12-5,FLUX-MET,202011171249
```