### 2020-12-30 OR_original_model.Rmd
Flushed out analysis of basic real-world application. Methods: model description

---
title: "OR Multi Disturbance"
author: "Michael Dietze"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("disturbance.R")
source("multi.ens.adj.R")
```

## set up
```{r}
## Load disturbance rate, state&transition, disturbed biomass stats
load("STprior.RData")
colnames(dextract) <- paste0("t",dtime)
dextract[is.na(dextract)] <- 0
d.prior["UN"] = d.prior["UN"] + d.prior[6]                         ## change at runtime (removes <NA>)
d.prior = d.prior[-6]

## Load Calibration
load("fit.RData")
Qsoil = Qs

## Load extracted site timeseries
load("extract.h03v04.RData")
```

## drivers
```{r}
latlon = points@coords
metFile = "daymet.RData"
if(file.exists(metFile)){
  load(metFile)
} else {
  dm = list()
  for(i in 1:nrow(latlon)){
    dm[[i]] = daymetr::download_daymet(site="h03v04",latlon[i,2],latlon[i,1],start=1985,end=2017,internal=TRUE,simplify=TRUE)
  }
  save(dm,file=metFile)
}

## extract Shortwave and convert to daily PAR
SW2PAR = 0.45
srad = subset(dm[[1]],measurement == "srad..W.m.2.")   
metTime <- as.POSIXct(paste(srad$year,srad$yday,sep="-"),format="%Y-%j")
rad <- matrix(NA,length(dm),length(metTime))
for(i in seq_along(dm)){
  srad = subset(dm[[i]],measurement == "srad..W.m.2.")   
  rad[i,] = srad$value*1e-6*86400*SW2PAR  # -> MJ /m2 /day PAR
}
plot(metTime,rad[1,],type='l')
## plot(date,PAR,col=2) ## calibration data
```

## functions
```{r,echo=FALSE,results='hide',message=FALS}
##' Single Site, Single Disturbance
##' @param Y observed data
##' @param R observation error
##' @param Q process error
##' @param Xic state initial condition matrix
##' @param THETA parameter matrix
##' @param dist  disturbance list: p.dist, mu0[3], V0[3,3], alloc soil 
SSSD.model <- function(Y,R,Q,PAR,Xic,THETA,dist){

  ## init
  NT = length(PAR)/365
  Nf = Nmu = Na = Xic
  ne = nrow(Xic)
  if(!is.matrix(THETA)) THETA = matrix(THETA,nrow=ne,ncol=length(THETA),byrow = TRUE)
  Forecast = Analysis = list()

  AnalysisSSSDMP <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      muN ~ dmnorm(mufN,pfN)                ## forecast prior on Not disturbed
      muD ~ dmnorm(mufD,pfD)                ## forecast prior on Disturbed
      D ~ dbern(p)                          ## disturbance probability
      ## Analysis 
      n <- D*muD + (1-D)*muN                ## select between disturbed and undisturbed
      Y[H] ~ dmnorm(n[H],R[H,H])                       ## Observation error
  }"

  for(t in 1:NT){

    ## update IC
    THETA[,9:11]  <- Na    ## Multinomial

    ## Forecast 
    for(i in 1:ne){
      tsel    = (t-1)*365 + 1:365             ## Select days from driver files
      Nmu[i,] = VSEM(THETA[i,], PAR = PAR[tsel])[365,2:4] ## deterministic simulation
      Nf[i,]  = mvtnorm::rmvnorm(1,Nmu[i,],Q) ## process error
    }

    ## Disturbance
    d.forecast <- rep(0,ne); d.forecast[1:ceiling(ne*0.1)] <- 1 #rbinom(ne,1,dist$p.dist)             ## prior on disturbance status
    sel <- !as.logical(d.forecast)                ## select undisturbed ensemble members
    if(sum(!sel)>2){
      sel.dist = which(!sel)
      for(j in sel.dist){                      ## run disturbance model
        Nf[j,]  = disturbance.p(Nf[j,],mu0=dist$mu0,V0 = dist$V0,alloc.soil = dist$alloc.soil)
      }
      mufD  = colMeans(Nf[!sel,])
      pfD = solve(var(Nf[!sel,]))#solve(diag(c(0.15,0.5,0.25)^2))
    } 
    if(sum(!sel) <= 2 || class(pfD) == "try-error") {   ## backup values if the ensemble stats fail
      mufD = c(dist$mu0[1], mean(Nf[,2]),  dist$mu0[2])            
      pfD  = solve(diag(c(0.15,0.5,0.25)^2))
    }
    Nf <- tobit(Nf)

    ## priors
    priors <- list(p    = dist$p.dist,         ## disturbance rate
                   mufN = colMeans(Nf[sel,]),    ## Not disturbed forecast means
                   mufD = mufD,   ## Disturbed forecast means
                   pfN  = solve(var(Nf[sel,])),  ## Not disturbed forecast precision
                   pfD  = pfD  ## Disturbed forecast precision
    )
    Forecast[[t]] <- c(list(d.forecast=d.forecast,Nf=Nf),priors)

    ## Analysis
    update = c(list(Y=Y[t,],R=solve(R)),priors,H = which(!is.na(Y[t,])))
    if(length(update$H) > 0){ ## data observed, run filter
      mod <- rjags::jags.model(file=textConnection(AnalysisSSSDMP),
                               data=update,
                               n.adapt=1000,n.chains=3,quiet=TRUE)
      jdat <- rjags::coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000)

      ## update parameters & save Analysis posterior
      dat = as.matrix(jdat)
      nsel <- grep(pattern="^n",x=colnames(dat))
      Xa = dat[,nsel]
      ca = dat[,"D"]
      run.adjustment = TRUE
      if(FALSE){
        d3 <-density(Xa[,3]) 
        plot(d3)
        lines(d3$x,dnorm(d3$x,update$Y[3],1/sqrt(R[3,3])),col=2)
        lines(d3$x,dnorm(d3$x,priors$mufN[3],1/sqrt(priors$pfN[3,3])),col="blue")
        p.kf = R[3,3]+priors$pfN[3,3]
        mu.kf = priors$mufN[3]*priors$pfN[3,3]/p.kf+
          update$Y[3]*R[3,3]/p.kf
        lines(d3$x,dnorm(d3$x,mu.kf,1/sqrt(p.kf)),col=3,lty=2)

        jtest <- "model{
         Y ~ dmnorm(n[H],R[H,H])
        }"
        jtest <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      n[3] ~ dnorm(mufN[3],pfN[3,3])  
      Y[H] ~ dnorm(n[H],R[H,H])
      dummy <- mufD[1] + pfD[1,1] + p
  }"
      jtest <- "  ## SingleSite SingleDisturbance MultiPool
    model{
      ## prior
      n[3] ~ dnorm(mufN[3],pfN[3,3])  
      Y[H] ~ dnorm(n[3],R[H,H])
  }"
        update2 = list(mufN = update$mufN,
                       pfN = update$pfN,
                       Y=update$Y,R=R,H=3)
        mtest <- rjags::jags.model(file=textConnection(jtest),
                                   data = update,
#                               data=list(n = c(NA,NA,mu.kf),
#                                         R=matrix(p.kf,3,3),H=3),
                               n.adapt=1000,n.chains=3,quiet=TRUE)
        mdat <- rjags::coda.samples(mtest,variable.names=c("n"),n.iter=10000) 
      summary(mdat)  
      }
    } else {
      Xa = Nf
      ca = 1
      run.adjustment = FALSE
    }

      ## draw new IC
    if(run.adjustment){
    Na <- multi.ens.adj(
      Xf   = Nf,
      cf   = d.forecast,
      mu.f = rbind(priors$mufN,priors$mufD),
      Pf   = abind::abind(solve(priors$pfN),
                            solve(priors$pfD),
                            along=3),
      Xa = Xa,
      ca = ca)
    } else {
      Na = cbind(rep(0,nrow(Nf)),Nf)
    }
    Ca <- Na[,1] ## updated classf assignments
    Na <- tobit(Na[,-1])
    #  sel.ic <- sample.int(nrow(dat), ne, replace=TRUE)
    #  Na <- dat[sel.ic,nsel]
    Analysis[[t]] <- list(Na = Na,
                          Ca = Ca,
                          CI = apply(dat,2,quantile,c(0.025,0.25,0.5,0.75,0.975)),
                          mu = apply(dat,2,mean,na.rm=TRUE)
    )
  }
  return(list(Forecast=Forecast,Analysis=Analysis))
}
```

## initial conditions & observations
```{r}
## BIOMASS
B2C = 0.5 ##biomass to carbon ratio
Yw <- bextract*B2C*1000*1e-4 ## convert Mg/ha  to  kg C 
Rw <- (84*B2C*1000*1e-4)^2 ## Observation Error: Kennedy et al 2018 Environ. Res. Lett. 13 025004; Fig 2 RMSE
## units: Mg/ha -> kg/m2

## grab the distribution of AGB for the first time point from all pixels
AGBprior <- Yw[,1]
ne = length(AGBprior) ## number of ensemble members

## spin up
PARlong <- as.vector(t(rad))
spin = VSEM(theta,PARlong)
thin = seq(nrow(spin)/2,nrow(spin),length=ne)
spinIC = spin[thin,2:4]
#for(i in 2:4){
#  hist(spin[thin,i],type='l')
#}

## disturbance recovery trajectory
nNBG = 200 ## years
NBG <- array(NA,c(ne,nNBG,3))
for(i in 1:ne){
  param = theta
  param[9:11] <- spinIC[i,]*c(0.01,1,0.01) ## set leaf and wood to 1%
  raw = VSEM(param,PARlong[1:(nNBG*365.25)])
  NBG[i,,] = raw[round(seq(365,nNBG*365.25,length=nNBG)),2:4]  ## grab pools at the end of each year
}

## for each AGBprior, grab the matching leaf and soil
ICprior <- matrix(NA,ne,3)
age = rep(NA,ne)
for(i in 1:ne){
  age[i] <- which.min((AGBprior[i]-NBG[i,,3])^2)  
  ICprior[i,] <- c(NBG[i,age[i],1:2],AGBprior[i])
}
hist(age)
colnames(ICprior) <- c("Cv","Cs","CR") 
pairs(ICprior)

```

## analyze sites - just one disturbance
```{r}
Rmult = c(1,0.1,0.01,0.001)
Qmult = c(1,10)
Pmult = c(1)
Ri = 4 #4
Qi = 1 #2
Pi = 1
#i = 181  # 177:263 = #fire
od = c(83:176,264:356)
for(i in 83:356){
  ofile = paste0("OR_SD_",
                 formatC(i, width = 3, format = "d", flag = "0"),
                 "_P",Pi,"Q",Qi,"R",Ri,".RData")
  print(ofile)
  NT = floor(ncol(rad)/365)
  Yi = matrix(NA,NT,3)  ## observed data
  Yi[1:ncol(Yw),3] = Yw[i,]  ## only Cw observed
  R = diag(rep(Rw,3))*Rmult[Ri]    ## observation error
  Q = diag(c(Ql,Qsoil,Qw)^2)*Qmult[Qi]  ## process error
  dtype = which(names(d.prior) == dextract[i,"t2004"])
  if(length(dtype)==0) break
  dist <- list(              ## disturbance parameters
    p.dist = d.prior[dtype]*Pmult[Pi],     ## disturbance probability
    mu0 = c(0.2,PBBAR[4,dtype]),   ## % retention [leaf,agb], first parameter made up, should include soil
    V0 =  diag(c(0.15,PBVAR[4,dtype])^2), #variance in retention
    alloc.soil = 0.05        ## made up  
  )
  test = SSSD.model(Y = Yi,R=R,Q=Q,
                  PAR=rad[i,],Xic=ICprior,THETA=theta[1:11],
                  dist = dist)
  save(test,file=ofile)
}
```

## visualization
```{r}
i = 181
pool.names=c("Cleaf","Csoil","Cstem")
for(i in 1:3){

    CI.Na = sapply(X = test$Analysis,function(x){
      y = x$CI
      sel.n = which(colnames(y) == paste0("n[",i,"]"))
      y[,sel.n]
    })

    CI.D = sapply(X = test$Analysis,function(x){
      y = x$CI
      sel.n = which(colnames(y) == paste0("D"))
      y[,sel.n]
    })

    CI.Nf = sapply(X = test$Forecast, function(x){
      mu = x$mufN[i]
      sd = 1/sqrt(x$pfN[i,i])
      return(mu + c(-1.96,0,1.96)*sd)
    })

  plot(Yi[,i],
       ylim = range(rbind(CI.Na,Yi[,i]),na.rm=TRUE),
       type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.Nf),CI.Nf[1,],CI.Nf[3,],col=ecoforecastR::col.alpha("blue",0.5))
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[5,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.Na[3,],col="#b2df8a",lwd=2)
  points(Yi[,i])
  lines(CI.D[3,])

}

```

 TODO:
 * why is Analysis not contracting around the data?
    * reducing obs error by /10 didn't solve
 * why are the disturbance forecast pools crazy?
 * assess disturbance posterior

## Computer experiments
* Run under defaults
* Run under a grid of inc process error and decrease obs error and increase disturbance probability

## postprocessing

```{r}
## Detect set of experiments that have been run
runs <- dir(pattern="*.RData")
runs <- runs[grepl("^OR",runs)]
exp = sapply(strsplit(runs,"_",fixed=TRUE),function(x){
  sub(".RData","",x[4],fixed = TRUE)
})
exps <- unique(exp)

## load experiments that haven't yet been summarized
# elist <- list()   ## run once
t04 = 15            ## time of disturbance assessement (2004)
tlo = 28            ## time of last observation
for(i in seq_along(exps)){
  if(is.null(elist[[exps[i]]])){
    esel <- which(exp==exps[i])
    distrate = rep(NA,nrow(dextract))
    divergence = predist = distsize = matrix(NA,nrow(dextract),6) ## pre and post disturbance biomass pools (w/uncertainty)
    for(j in seq_along(esel)){
      ## load file
      load(runs[esel[j]])
      k = as.numeric(unlist(strsplit(runs[esel[j]],"_",fixed=TRUE))[3])

      ## extract detection rate
      distrate[k]=test$Analysis[[t04]]$mu["D"]

      ## capture of disturbance change
      predist[k,1]    = test$Analysis[[t04-1]]$mu["n[3]"]
      predist[k,2:6]  = test$Analysis[[t04-1]]$CI[,"n[3]"]
      distsize[k,1]   = test$Analysis[[t04]]$mu["n[3]"]
      distsize[k,2:6] = test$Analysis[[t04]]$CI[,"n[3]"]

      ## Filter divergence
      divergence[k,1]   = test$Analysis[[tlo]]$mu["n[3]"]
      divergence[k,2:6] = test$Analysis[[tlo]]$CI[,"n[3]"]
    } ## end loop over individual runs

    elist[[exps[i]]] = list(distrate=distrate,predist=predist,
                            distsize=distsize,divergence=divergence)

  } ## end if not processed

} ## end loop over experiments

## grab matching data
obs.pre  = Yw[,t04-1]
obs.dist = Yw[,t04]
obs.tlo  = Yw[,tlo]
obs.class= dextract[,"t2004"]/10
```

## analyses

```{r}
pdf("OR_exps_diagnostics.pdf")
trim <- function(x,lo=0.0001,hi=0.9999){
  x[x<lo] <- lo
  x[x>hi] <- hi
  return(x)
}
betatrim <- function(y){
    n.obs <- sum(!is.na(y))
    (y * (n.obs - 1) + 0.5) / n.obs
}
stats <- list()
for(i in seq_along(exps)){

  ## divergence
  div = (elist[[i]]$divergence[,1] - obs.tlo)/
    (apply(elist[[i]]$divergence[,c(2,6)],1,diff)*0.5)
  div.reg    = lm(div~obs.tlo)
  div.median = median(div,na.rm = TRUE)
  div.cov    = sum(div < 1 & div > -1,na.rm=TRUE)/sum(!is.na(div))

  hist(div,main=exps[i])
  abline(v=c(-1,1),lty=2)
  plot(obs.tlo,div,col=obs.class,main=exps[i])
  abline(h=c(-1,1),lty=2)
  abline(div.reg,col=2)

  ## Disturbance detection
  distrate = elist[[i]]$distrate
  hist(distrate,main=exps[i])

  ## Disturbance skill
  ord <- order(distrate)
  plot(distrate[ord],col=obs.class[ord],main=exps[i])
  dist.bar   = mean(distrate,na.rm = TRUE)
  dist.class = tapply(distrate,obs.class,mean,na.rm=TRUE)

  ## vs predisturbance biomass
  distrate2 = betatrim(distrate)
  pred.seq = data.frame(obs.pre=seq(0,max(obs.pre,na.rm = TRUE),length=500),
                        distprop = seq(-1,1,length=500))
  dist.reg.pre  = lm(distrate ~ obs.pre)
  dist.beta.pre = betareg::betareg(distrate2 ~ obs.pre,
                                   na.action=na.omit,link="logit")
  plot(obs.pre,distrate,col=obs.class,main=exps[i])
  abline(dist.reg.pre,col=2)
  lines(pred.seq$obs.pre,predict(dist.beta.pre,newdata=pred.seq),col="green")

  ## vs proportional disturbance size
  distprop = 1 - elist[[i]]$distsize[,1]/elist[[i]]$predist[,1] ## predicted
  distprop = 1 - obs.dist/obs.pre ## observed
  dist.beta.prop = betareg::betareg(distrate2 ~ distprop,
                                   na.action=na.omit,link="logit")
  plot(distprop,distrate,col=obs.class,main=exps[i])
  lines(pred.seq$distprop,predict(dist.beta.prop,newdata=pred.seq),col="green")

  ## vs ABSOLUTE disturbance size
  distabs = obs.pre-obs.dist
  pred.seq$distabs = seq(min(distabs),max(distabs),length=500)
  dist.beta.abs = betareg::betareg(distrate2 ~ distabs | distabs,
                                   na.action=na.omit,
                                   link="logit",
                                   link.phi = "identity",
                                   start=c(0,5,0.5,0),
                                   type="BR")
  dist.qb.abs = glm(distrate ~ distabs,
                    na.action=na.omit,
                    family=quasibinomial("logit")
  )
  dist.bin.abs = glm(round(distrate) ~ distabs,
                    na.action=na.omit,
                    family=binomial("logit")
  )
  plot(distabs,distrate,col=obs.class,main=exps[i])
  lines(pred.seq$distabs,predict(dist.beta.abs,newdata=pred.seq,type="quantile"),col="green")
  lines(pred.seq$distabs,predict(dist.beta.abs,newdata=pred.seq,type="quantile",at=0.75),col="green",lty=2)
  lines(pred.seq$distabs,predict(dist.beta.abs,newdata=pred.seq,type="quantile",at=0.25),col="green",lty=2)
  lines(pred.seq$distabs,predict(dist.qb.abs,newdata=pred.seq,type="response"),col="purple")
  lines(pred.seq$distabs,predict(dist.bin.abs,newdata=pred.seq,type="response"),col="blue")

  ## vs divergence
  absdiv = abs(div)
  dist.beta.absdiv = betareg::betareg(distrate2 ~ absdiv,
                                   na.action=na.omit,link="logit")
  dist.bin.absdiv = glm(round(distrate) ~ absdiv,
                                   na.action=na.omit,
                        family = binomial("logit"))
  pred.seq$absdiv = seq(0,max(absdiv,na.rm=TRUE),length=500)
  plot(absdiv,distrate,col=obs.class,main=exps[i])
  abline(v=c(-1,1),lty=2,col=3)
  abline(lm(distrate ~ absdiv),col=2)
  lines(pred.seq$absdiv,predict(dist.beta.absdiv,newdata=pred.seq),col="green")
  lines(pred.seq$absdiv,predict(dist.bin.absdiv,newdata=pred.seq,type="response"),col="blue")

 dist.qb.combo = glm(distrate2 ~ obs.pre + distprop +
                        distabs + absdiv + obs.class,
                      na.action=na.omit,
                      family=quasibinomial("logit"))
  dist.bin.combo = glm(round(distrate) ~ obs.pre + distprop +
                        distabs + absdiv + as.factor(obs.class),
                      na.action=na.omit,
                      family=binomial("logit"))

# dist.beta.combo = betareg::betareg(distrate2 ~ obs.pre + obs.class + distabs + absdiv,
#                                   na.action=na.omit,link="logit")

  stats[[i]] <- list(
    div.reg = div.reg,
    div.median = div.median,
    div.cov = div.cov,
    dist.bar = dist.bar,
    dist.class = dist.class,
    dist.pre =  dist.beta.pre,
    dist.prop = dist.beta.prop,
    dist.beta.abs =  dist.beta.abs,
    dist.bin.abs  = dist.bin.abs,
    dist.absdiv   = dist.beta.absdiv,
    dist.combo = dist.bin.combo
  )

}
dev.off()
```

factors affecting detection of disturbance
```{r}
Ps = as.numeric(substr(names(elist),2,2))
Qs = as.numeric(substr(names(elist),4,4))
Rs = as.numeric(substr(names(elist),6,6))

## P

## Q

## R

## Divergence
div.cov = sapply(stats,"[[","div.cov")
plot(Rmult[Rs],div.cov,col=Qs,log="x",xlab="Variance Multiplier",ylab="Filter Coverage")

div.median = sapply(stats,"[[","div.median")
plot(Rmult[Rs],div.median,col=Qs,log="x",xlab="Variance Multiplier",ylab="Median Divergance")

div.slope = sapply(stats,function(x){coef(x$div.reg)[2]})
plot(Rmult[Rs],div.slope,col=Qs,log="x",xlab="Variance Multiplier",ylab="Divergance Slope")
## Disturbance @ default
summary(stats[[4]]$dist.combo)
## nothing matters

## Disturbance @ P1Q2R2 (intermediate)
summary(stats[[4]]$dist.combo)
## detection varies with divergence & disturbance type

## Disturbance magnitude: P1Q1R4
i = 3
summary(stats[[i]]$dist.combo)  ## absolute magnitude of divergence is best predictor
summary(stats[[i]]$dist.bin.abs)
beta = coef(stats[[i]]$dist.bin.abs)
## threshold mx+b = 0
thresh = -beta[1]/beta[2]

## Disturbance type: P1Q1R4 (reduced divergence)
stats[[i]]$dist.class
tapply(distabs,obs.class,summary) ## obs dist size by class
thresh.prop = tapply(distabs,obs.class,function(x){sum(x>thresh,na.rm = TRUE)/sum(!is.na(x))}) ## prop > detection threshold
thresh.prop
stats[[i]]$dist.class
dclass = stats[[i]]$dist.class
#dclass[1] = 0
dcol = as.numeric(names(stats[[i]]$dist.class)) + 1
plot(thresh.prop,dclass,col=dcol,pch=19)
abline(0,1,lty=2,col=2)
legend("topleft",legend = dcol-1,col=dcol,pch=19)

## additional experiments to run (P1Q1R4)
```

false positives

false negatives

classification success

## Next steps:
* Run with multiple disturbance options and PFT switching