### 2021-06-10 Paper.Rmd
model output for EnKF figure; config to render paper to PDF

---
title: "Disturbance Assimilation"
author: "Michael Dietze"
date: '`r Sys.Date()`'
output:
  pdf_document: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    keep_tex: true
  html_document: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    keep_tex: true
bibliography: bibliography.bibtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract {-}

# Intro

**Borrow text from CMS proposals**

1. The context: MRV and predicting Terrestrial C
2. Status quo: LULC * look up table
  + very coarse, doesn't capture spatiotemporal varibility
  + rightly captures that disturbance and land use change both critical to C accounting and, more importantly, a place where we can exert some control
  + BULC
  + state-and-transition modeling
  + process-model projections
3. Alternative: SDA
  + models capture process
  + data constrains (data driven, model as cov)
  + data fusion
  + seamless transition from reanalysis to nowcast to forecast [@Dietze2017a]
  + Disturbance is exactly why you would want to nudge a model back to reality
4. The problem: Conventional SDA doesn't capture jumps
  + series of nudges
  + potentially misattributes pools and fluxes
  + "drags down" neighbors **no sims of this yet**
5. The solution:
  + assimilate P(dist)
  + update C | dist
6. This paper
  + Illustration of problem
  + Derivation Multinomial
  + Simulated data experiments (known disturbance)
  + Real-world example

<!---- METHODS --->
```{r child = 'methods.Rmd'}
```

<!---- RESULTS --->
```{r child = 'results.Rmd'}
```

<!---- DISCUSSION --->
```{r child = 'discussion.Rmd'}
```

# Acknowledgements

NASA CMS, PEcAn, NEFI

# Bibliography