### 2021-06-14 TwoSite_SingleDisturb_MultiPool.Rmd
Results: exp 5 two site plot

---
title: "Two Site, Single Disturbance, MultiPool"
author: "Mike Dietze"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BayesianTools)
library(magrittr)
set.seed(42)
## helper functions
source("multi.ens.adj.R")
source("disturbance.R")
```

## SET-UP

```{r}
NT = 25
nt = 365*NT
time = 1:nt

## drivers
nd   = 31 ## number of driver ensembles
PAR <- matrix(NA,nd,nt)
for(i in 1:nd){
  PAR[i,] <- VSEMcreatePAR(time)  
}
plot(PAR[1,], main = "PAR (driving the model)", xlab = "Day",cex=0.35)

## Parameters
refPars <- VSEMgetDefaults()
refPars$lower[9:11] = 1
np = 50
sigmaP = diag((refPars$upper - refPars$lower)/10)^2 ## assume range is +/- 5 SD
Theta <- as.data.frame(mvtnorm::rmvnorm(np,refPars$best,sigmaP))
colnames(Theta) <- rownames(refPars)
par(mfrow=c(3,4))
for(i in 1:11){
  Theta[Theta[,i] < refPars$lower[i] , i] <- refPars$lower[i]
  Theta[Theta[,i] > refPars$upper[i] , i] <- refPars$upper[i]
  hist(Theta[,i],main=colnames(Theta)[i])
}

## initial conditions
## spin up
ne = 1000
met = sample.int(nd,ne,replace=TRUE)
par = sample.int(np,ne,replace=TRUE)
Na <- matrix(NA,ne,3)
for(i in 1:ne){
  Na[i,] = VSEM(as.numeric(Theta[par[i],]),PAR = PAR[met[i],])[nt,2:4]
}
Na.orig = Na
par(mfrow=c(1,3))
hist(Na[,1],main="Cleaf")
hist(Na[,2],main="Csoil")
hist(Na[,3],main="Cstem")
```

## set up storage and parameters

```{r}
## hard coded 
Q = diag(c(0.1,0.1,0.1)) ## process error
R = diag(c(.5,10,1))      ## observation error
R = as.matrix(Matrix::bdiag(R,R))  ## nsite = 2
THETA <- as.matrix(Theta[par,]) ## ensemble parameters
THETA = abind::abind(THETA,THETA,along = 3)  ## n=2 sites
mu0 <- c(1,1)            ## mean leaf & stem biomass after disturbance
V0  <- diag(c(0.15,0.25))^2## var in post disturbance leaf and stem biomass
alloc.soil = 0.25        ## fraction of removed C that goes in the soil
p.dist = 0.025

## storage
Dbar = 0
Analysis <- Forecast <- list()
Forecast.KD = Analysis.KD = list() ## EnKF with Disturbance

## pseudodata
nsite = 2
No = matrix(NA,NT,3*nsite)                   ## storage for observed state
Ntrue = matrix(NA,NT,3*nsite)                ## storage for latent "true" state
true = 42                                    ## reference ensemble member
D.true = matrix(0,nsite,NT)                  ## true disturbance timeseries
D.true[1,10] = 1                             ## prescribe a disturbance
refpar = matrix(as.numeric(Theta[par[true],]),nsite,ncol(Theta),byrow = TRUE) ## grab the "true" parameters
refpar[,9:11] = Na[true+(1:nsite)-1,]        ## set initial biomass pools
ref.f = list()
for(t in 1:NT){
  tsel <- (t-1)*365 + 1:365                  ## select 1 yr of met  
  for(s in 1:nsite){
    ref.mu = VSEM(refpar[s,],PAR[met[true],tsel])[365,2:4]    ## forward simulation
    ref.f[[s]]  = mvtnorm::rmvnorm(1,ref.mu,Q)## process error
    if(D.true[s,t]){
     ref.f[[s]] = disturbance(ref.f[[s]])
    }
    ref.f[[s]] = tobit(ref.f[[s]])
    refpar[s,9:11] = ref.f[[s]]                 ## update IC for next iteration
  }
  Ntrue[t,] = unlist(ref.f)
  No[t,] = tobit(mvtnorm::rmvnorm(1,Ntrue[t,],R)) ## observation error
}
par(mfrow=c(3,1))
plot(No[,1])
lines(No[,4],col=2)
plot(No[,2])
lines(No[,5],col=2)
plot(No[,3])
lines(No[,6],col=2)
```

```{r,echo=FALSE,results='hide',message=FALSE}
Analysis2SMDMP <- "  ## SingleSite MultiDisturbance MultiPool
model{
  ## prior
  for(i in 1:N){
    x[1:6,i]  ~ dmnorm(muf[,i],pf[,,i]) ## forecast priors for different cases
  }
 rho[2] ~ dbern(p[2])
 rho[3] ~ dbin(p[3]/(1-p[2]),1-rho[2])
 rho[4] ~ dbin(p[4]/(1-p[2]-p[3]),1-rho[2]-rho[3])
 rho[1] <- 1 - rho[2] - rho[3] - rho[4]
  X <- rho[1]*x[1:6,1] + rho[2]*x[1:6,2] + rho[3]*x[1:6,3] + rho[4]*x[1:6,4] ## select between cases
  Y ~ dmnorm(X,R)                       ## Observation error
}"

Na <- NaK <- abind::abind(Na.orig,
                          Na.orig[sample.int(ne,ne),],  ## resample IC for site 2
                          along = 3)
NfK = Nf = Na*NA
THETAK <- THETA
d.prior = matrix(nrow=ne,ncol=2)
muf <- matrix(NA,3*nsite,4)
pf  <- array(0,c(3*nsite,3*nsite,4)) ## 3rd variable is class
for(t in 1:NT){

  tsel <- (t-1)*365 + 1:365               ## Select days from driver files

  ## update IC
  THETA[,9:11,]  <- Na    ## Multinomial
  THETAK[,9:11,] <- NaK  ## EnKF w/ disturbance

  ## Forecast 
  for(i in 1:ne){
    for(s in 1:nsite){
      Nmu  =  VSEM(THETA[i,,s], PAR = PAR[met[i],tsel])[365,2:4]    
      NmuK      = VSEM(THETAK[i,,s],PAR = PAR[met[i],tsel])[365,2:4] ## deterministic simulation
      epsilon  = mvtnorm::rmvnorm(1,rep(0,3),Q)                   ## process error: shared between versions, currently independent by site
      Nf[i,,s]   = Nmu  + epsilon
      NfK[i,,s]  = NmuK + epsilon
    }
  }

  ## Disturbance
  for(i in 1:ne) d.prior[i,] <- rcbern(rep(p.dist,2),rho=0.5,mu=1:2)   ## prior on disturbance status
  d.class <- d.prior[,1] + 2*d.prior[,2]
  for(d in 2:4){  ## loop over disturbance classes
    sel.dist <- which(d.class == d-1)              ## select disturbed ensemble members
    if(length(sel.dist)> (3*nsite)){
      for(j in sel.dist){                      ## run disturbance model
        switch (as.character(d),
                "2" = {Nf[j,,1] = disturbance(Nf[j,,1])
                NfK[j,,1]= disturbance(NfK[j,,1])
                },
                "3" = {Nf[j,,2] = disturbance(Nf[j,,2])
                NfK[j,,2]= disturbance(NfK[j,,2])
                },
                "4" = {Nf[j,,1] = disturbance(Nf[j,,1])
                Nf[j,,2] = disturbance(Nf[j,,2])
                NfK[j,,1]= disturbance(NfK[j,,1])
                NfK[j,,2]= disturbance(NfK[j,,2])
                }
        )
      }
      muf[,d]  = apply(Nf[sel.dist,,],3,colMeans)
      pfTMP    = try(solve(var(cbind(Nf[sel.dist,,1],Nf[sel.dist,,2]))))
    } 
    if(sum(sel.dist) <= (3*nsite) || class(pfTMP) == "try-error") { 
      ## set to highly unlikely values that will be selected out
      muf[,d]  = rep(0,nsite*3)
      pfTMP    = solve(diag(rep(0.01,nsite*3)^2))
    }
    pf[,,d]    = pfTMP
  }
  ## set undisturbed
  sel.dist = which(d.class == 0)
  muf[,1] = apply(Nf[sel.dist,,],3,colMeans)
  pf[,,1] = solve(var(cbind(Nf[sel.dist,,1],Nf[sel.dist,,2])))
  Nf <- tobit(Nf)

  ## EnKF forecast stats
  if(FALSE){
    NfK <- tobit(NfK)
    NK = cbind(NfK[,,1],NfK[,,2])
    mufK = colMeans(NK)
    PfK  = var(NK)
    pfK  = solve(PfK)             ## Not disturbed forecast precision
    Forecast.KD[[t]] <- list(d.prior=d.prior,Nf=NfK,muf=mufK,pf=pfK)
  }
  ## priors
  priors <- list(N = 4,  ## 2 sites x 2 disturbance classes
                 p   = table(d.class)/ne,         ## disturbance rate
                 muf = muf,    ## forecast means
                 pf  = pf      ## forecast precision
  )
  Forecast[[t]] <- c(list(d.prior=d.prior,Nf=Nf),priors)

  ## initial conditions: disturbance likelihood
  Lp = NA
  for(i in 1:4){ Lp[i] = mvtnorm::dmvnorm(No[t,],muf[,i],pf[,,i])}
  rho.IC = rep(0,4); rho.IC[which.max(Lp)] = 1; rho.IC[1] = NA

  ## Analysis
  update = c(list(Y=No[t,],R=solve(R)),priors)
  mod <- rjags::jags.model(file=textConnection(Analysis2SMDMP),
                           data=update,
                           inits = list(rho=rho.IC),
                           n.adapt=1000,n.chains=3)
  jdat <- rjags::coda.samples(mod,variable.names="rho",n.iter=2000) ## burn-in 
  jdat <- rjags::coda.samples(mod,variable.names=c("X","rho","x"),n.iter=10000)

  ## update parameters & save Analysis posterior
  dat = as.matrix(jdat)
  Xsel <- grep(pattern="^X",x=colnames(dat))

  ## draw new IC
  d.post <- dat[,grep(pattern="rho",x=colnames(dat))]
  d.class.post <- apply(d.post,1,which.max)
  Na <- multi.ens.adj(
    Xf   = cbind(Nf[,,1],Nf[,,2]),
    cf   = d.class+1,
    mu.f = t(priors$muf),
    Pf   = abind::abind(    ## find a way to do this with apply(priors$pf,3,solve)
      solve(priors$pf[,,1]),
      solve(priors$pf[,,2]),
      solve(priors$pf[,,3]),
      solve(priors$pf[,,4]),
      along=3),
    Xa = dat[,Xsel],
    ca = d.class.post)
  Ca <- Na[,1] ## updated class assignments
  Na <- tobit(Na[,-1])
  #  sel.ic <- sample.int(nrow(dat), ne, replace=TRUE)
  #  Na <- dat[sel.ic,nsel]
  Analysis[[t]] <- list(MCMC=dat,
                        Na = Na,
                        Ca = Ca)
  Na <- abind::abind(Na[,1:3],Na[,4:6],along = 3)

  ## EnKF Analysis
  if(FALSE){
    I = H <- diag(rep(1,3*nsite))
    K <- PfK %*% t(H) %*% solve(H%*%PfK%*%t(H) + R)
    muaK <- mufK + K%*%(No[t,] - mufK) 
    PaK <- (I-K %*% H)%*%PfK    
    paK = solve(PaK)
    NaK <- PEcAn.assim.sequential::adj.ens(Pf = PfK,X = NK,mu.f = mufK,mu.a = muaK,Pa = PaK) %>% as.matrix()
    Analysis.KD[[t]] <- list(mua=muaK,pa=paK,Pa=PaK,Na=NaK)
    NaK <- abind::abind(NaK[,1:3],NaK[,4:6],along = 3)
  }

}

```

```{r}
save(Forecast,Analysis,No,Ntrue,R,file="exp5.RData")
Forecast10 = Forecast[[10]]
Analysis10 = Analysis[[10]]
CI.Na = list()
for(i in c(3,6)){
  CI.Na[[i]] = sapply(X = Analysis,function(y){
    x = y$MCMC
    sel.n = which(colnames(x) == paste0("X[",i,"]"))
    quantile(x[,sel.n],c(0.025,0.5,0.975))
  })
}
save(Forecast10,Analysis10,CI.Na,No,Ntrue,R,file="exp5trim.RData")

```

### time series

```{r}
pool.names=c("Cleaf","Csoil","Cstem")
for(i in 1:6){

  CI.Na = sapply(X = Analysis,function(y){
    x = y$MCMC
    sel.n = which(colnames(x) == paste0("X[",i,"]"))
    quantile(x[,sel.n],c(0.025,0.5,0.975))
  })
  #CI.Na = sapply(X = Analysis,function(x){
  #  apply(x,2,quantile,c(0.025,0.975))
  #})
  CI.EnKF = sapply(Analysis.KD,function(x){
    mu = x$mua[i]
    sd = sqrt(x$Pa[i,i])
    return(c(mu-1.96*sd,mu,mu+1.96*sd))
  })

  plot(No[,i],
       #ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
       ylab=pool.names[i],xlab="time")
  ecoforecastR::ciEnvelope(1:ncol(CI.EnKF),CI.EnKF[1,],CI.EnKF[3,],col=ecoforecastR::col.alpha('#b2df8a',0.7))
  lines(CI.EnKF[2,],col="#b2df8a",lwd=2)
  ecoforecastR::ciEnvelope(1:ncol(CI.Na),CI.Na[1,],CI.Na[3,],col=ecoforecastR::col.alpha('blue',0.5))
  lines(CI.Na[2,],col="blue",lwd=2)
  abline(v=10)  
  lines(Ntrue[,i],lty=2,lwd=2)
}
```
### Ensemble adj at time of disturbance

```{r,fig.asp=1}
#par(mfrow=c(1,3))

plot(Analysis[[10]]$Na[,1],Analysis[[10]]$Na[,3],pch="+",col=ecoforecastR::col.alpha("red",alpha=0.25),
     ylim=range(Forecast[[10]]$Nf[,3,1]),
     xlim=range(Forecast[[10]]$Nf[,1,1]),
     xlab=pool.names[1],ylab=pool.names[3],
     cex.lab=1.25,asp = 1)
points(Forecast[[10]]$Nf[,1,1],Forecast[[10]]$Nf[,3,1],cex=0.5,col=ecoforecastR::col.alpha("black",alpha=0.25))
points(No[10,1],No[10,3],cex=3,col=ecoforecastR::col.alpha("purple",alpha = 0.75),pch=18)

## ensemble adjustment
sel.a=sample.int(ne,10)
arrows(Forecast[[10]]$Nf[sel.a,1,1],Forecast[[10]]$Nf[sel.a,3,1],
       Analysis[[10]]$Na[sel.a,1],Analysis[[10]]$Na[sel.a,3],length = 0.1,lwd = 0.5)   
legend("topleft",
       legend=c("Forecast","Analysis","Data"),
       pch=c(1,3,18),col=c(1,2,"purple"),
       cex = 1.5)

cov2cor(solve(Forecast.KD[[10]]$pf))

```
## EnKF, no disturbance

```{r,echo=FALSE}
Na = Na.orig
Forecast.EnKF = Analysis.EnKF = list()
for(t in 1:NT){

  tsel <- (t-1)*365 + 1:365               ## Select days from driver files

  ## update IC
  THETA[,9:11] <- Na

  ## Forecast 
  for(i in 1:ne){
    Nmu[i,] = VSEM(THETA[i,],PAR = PAR[met[i],tsel])[365,2:4] ## deterministic simulation
    Nf[i,]  = mvtnorm::rmvnorm(1,Nmu[i,],Q) ## process error
  }

  muf = colMeans(Nf)    ## Not disturbed forecast means
  Pf  = var(Nf)
  pf  = solve(Pf)             ## Not disturbed forecast precision
  Forecast.EnKF[[t]] <- list(d.prior=d.prior,Nf=Nf,muf=muf,pf=pf)

  ## Analysis
  I = H <- diag(rep(1,3))
  K <- Pf %*% t(H) %*% solve(H%*%Pf%*%t(H) + R)
  mua <- muf + K%*%(No[t,] - muf) 
  Pa <- (I-K %*% H)%*%Pf    
  pa = solve(Pa)

  ## draw new IC
  #  Na <- mvtnorm::rmvnorm(ne,mua,Pa)
  Na <- PEcAn.assim.sequential::adj.ens(
    Pf = Pf,X = Nf,mu.f = muf,mu.a = mua,Pa = Pa)

  ## save Analysis posterior
  Analysis.EnKF[[t]] <- list(mua=mua,pa=pa,Pa=Pa,Na=Na)

}

```