### 2023-01-01 SingleSite_SingleDisturb.Rmd
Results: simulated data experiments

---
title: "Single Site, Single Disturbance"
author: "Michael Dietze"
date: "7/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(ecoforecastR)
library(tidyverse)
```

Initial code for this started from the "Assimilation of disturbance" example in <https://github.com/EcoForecast/EF_BookCode/blob/master/Ch02_Logitic.Rmd> which itself derived from simple simulations and KF developed for a 2013 talk at the Ecological Society of America annual meeting.

### Assimilation of disturbance

"True" run

```{r,echo=1:2}
#fig.width=7, fig.height=6
set.seed(1066)
p.dist = 0.005
n0 = 1
nIC = 6
r = .1
K = 6
n.sd = 0.25
NT = 100
time = 1:NT
xlim = c(5,25)
n = rep(nIC,NT)
d = rep(0,NT)
d[10] = 1  ## prescribe a disturbance
for(t in 2:NT){
  n[t] = n[t-1] + r*n[t-1]*(1-n[t-1]/K) + rnorm(1,0,n.sd)
#  d[t] = rbinom(1,1,p.dist)
  if(d[t]) n[t] = abs(rnorm(1,n0,0.5))
}

plot(time,n,type='l',ylim=range(n),xlim=xlim,ylab="LAI",lwd=3)
obs.sd = 0.75
n.true=n
n.obs = rnorm(NT,n.true,obs.sd)
points(time,n.obs,pch=18,col=2,cex=1.5)
legend("topright",legend=c("truth","obs"),lwd=c(2,NA),pch=c(NA,18),col=1:2)
## Simulate less frequent observations
#na.seq = 1:NT; na.seq=na.seq[-seq(1,NT,by=5)]
#n.fia = n.obs;n.fia[na.seq] <- NA
#points(time,n.fia,pch=18,col="blue",cex=2)
```

## "blind" forecast w/o disturbance

```{r}
n0.sd = n.sd
NE = 10000
n0s = rlnorm(NE,log(nIC),n0.sd)
ne0 = matrix(n0s,NE,NT)
for(i in 1:NE){
  for(t in 2:NT){
    ne0[i,t] = abs(ne0[i,t-1] + r*ne0[i,t-1]*(1-ne0[i,t-1]/K) + rnorm(1,0,n.sd))
#    if(rbinom(1,1,p.dist)) ne[i,t] = abs(rnorm(1,n0,0.5))
  }
}
yrng = c(0,8) #range(ne0)
n0.stats = apply(ne0,2,quantile,c(0.025,0.5,0.975))
plot(time,n.obs,type='n',ylim=yrng,xlim=xlim)
polygon(cbind(c(time, rev(time), 1), c(n0.stats[1, ], rev(n0.stats[3, ]),
  n0.stats[1, 1])), border = NA, col = 'lightblue')
lines(time,n0.stats[2,],lwd=2,col="blue4")
points(time,n.obs,pch=18,col=2,cex=1.5)
```

## standard EnKF w/o disturbance

```{r,echo=FALSE, fig.width=7, fig.height=6}
## initial ensemble
ens = matrix(NA,NT*2+1,NE)
ens[1,] = rlnorm(NE,log(nIC),n0.sd)
Xae = Pae = 0
Xfe = mean(ens[1,])
Pfe = var(ens[1,])
for(i in 1:NT){

  ##Approximate forecast with Normal
  k = 2*(i-1)+1
  Xfe[i] = mean(ens[k,])
  Pfe[i] = var(ens[k,])

  ## analysis step: combine previous forecast with observed data
  H <- matrix(1,ncol=1)
  Kg <- Pfe[i] * t(H) %*% solve(H%*%Pfe[i]%*%t(H) + obs.sd)
  if(!is.na(n.obs[i])){
    Xae[i] <- Xfe[i] + Kg*(n.obs[i] - Xfe[i]) 
  }else{
    Xae[i] <- Xfe[i]
  }
  Pae[i] <- (1-Kg %*% H)*Pfe[i]

  ## generate new ensemble
  ens[k+1,] = rnorm(NE,Xae[i],sqrt(Pae[i]))

  ## Forecast step: predict to next step from current
  ens[k+2,] = ens[k+1,] + r*ens[k+1,]*(1-ens[k+1,]/K) + rnorm(NE,0,n.sd) 
#  disturbed = as.logical(rbinom(NE,1,p.dist))
#  ens[k+2,disturbed] = abs(rnorm(sum(disturbed),n0,0.5)) ## disturbance

}

## forecast
fx0 = ens[2*(1:NT)-1,]
fx0.stats = apply(fx0,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,fx0.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)  
polygon(cbind(c(time, rev(time), 1), c(n0.stats[1, ], rev(n0.stats[3, ]),      ## Free run, no disturbance
  n0.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(fx0.stats[1, ], rev(fx0.stats[5, ]),    ## Forecast 95%
  fx0.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(fx0.stats[2, ], rev(fx0.stats[4, ]),    ## Forecast IQR
  fx0.stats[2, 1])), border = NA, col = 'green4')
#lines(time[-NT],fx.stats[2,-1],lwd=4,col="green")
lines(time,n0.stats[2,],lwd=2,col="blue4")                                     ## Free run mean
lines(time,n.true,type='l',lwd=2)                                              ## truth
points(time,n.obs,pch=18,col=2)                                                ##obs
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

## analysis
an0 = ens[2*(1:NT),]
an0.stats = apply(an0,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,an0.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)
polygon(cbind(c(time, rev(time), 1), c(n0.stats[1, ], rev(n0.stats[3, ]),
  n0.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(an0.stats[1, ], rev(an0.stats[5, ]),
  an0.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[2, 1])), border = NA, col = 'green4')
lines(time,n0.stats[2,],lwd=2,col="blue4")
lines(time,n.true,type='l',lwd=2)
points(time,n.obs,pch=18,col=2)
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

```

## "blind" forecast w/ disturb

```{r}
n0.sd = n.sd
NE = 10000
n0s = rlnorm(NE,log(nIC),n0.sd)
ne = matrix(n0s,NE,NT)
for(i in 1:NE){
  for(t in 2:NT){
    ne[i,t] = abs(ne[i,t-1] + r*ne[i,t-1]*(1-ne[i,t-1]/K) + rnorm(1,0,n.sd))
    if(rbinom(1,1,p.dist)) ne[i,t] = abs(rnorm(1,n0,0.5))
  }
}
yrng = range(ne)
n.stats = apply(ne,2,quantile,c(0.025,0.5,0.975))
plot(time,n.obs,type='n',xlim=xlim)
polygon(cbind(c(time, rev(time), 1), c(n.stats[1, ], rev(n.stats[3, ]),
  n.stats[1, 1])), border = NA, col = 'lightblue')
lines(time,n.stats[2,],lwd=2,col="blue4")
points(time,n.obs,pch=18,col=2,cex=1.5)
```

## standard EnKF + Disturb

```{r,echo=FALSE, fig.width=7, fig.height=6}
## initial ensemble
ens = matrix(NA,NT*2+1,NE)
ens[1,] = rlnorm(NE,log(nIC),n0.sd)
Xae = Pae = 0
Xfe = mean(ens[1,])
Pfe = var(ens[1,])
for(i in 1:NT){

  ##Approximate forecast with Normal
  k = 2*(i-1)+1
  Xfe[i] = mean(ens[k,])
  Pfe[i] = var(ens[k,])

  ## analysis step: combine previous forecast with observed data
  H <- matrix(1,ncol=1)
  Kg <- Pfe[i] * t(H) %*% solve(H%*%Pfe[i]%*%t(H) + obs.sd)
  if(!is.na(n.obs[i])){
    Xae[i] <- Xfe[i] + Kg*(n.obs[i] - Xfe[i]) 
  }else{
    Xae[i] <- Xfe[i]
  }
  Pae[i] <- (1-Kg %*% H)*Pfe[i]

  ## generate new ensemble
  ens[k+1,] = rnorm(NE,Xae[i],sqrt(Pae[i]))

  ## Forecast step: predict to next step from current
  ens[k+2,] = ens[k+1,] + r*ens[k+1,]*(1-ens[k+1,]/K) + rnorm(NE,0,n.sd) 
  disturbed = as.logical(rbinom(NE,1,p.dist))
  ens[k+2,disturbed] = abs(rnorm(sum(disturbed),n0,0.5)) ## disturbance
#  Xfe[i+1] = m*Xa[i] + beta[2]
#  Pfe[i+1] = q + m*Pa[i]*m

}

## forecast
fx = ens[2*(1:NT)-1,]
fx.stats = apply(fx,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,fx.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)
polygon(cbind(c(time, rev(time), 1), c(n.stats[1, ], rev(n.stats[3, ]),
  n.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(fx.stats[1, ], rev(fx.stats[5, ]),
  fx.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(fx.stats[2, ], rev(fx.stats[4, ]),
  fx.stats[2, 1])), border = NA, col = 'green4')
#lines(time[-NT],fx.stats[2,-1],lwd=4,col="green")
lines(time,n.stats[2,],lwd=2,col="blue4")
lines(time,n.true,type='l',lwd=2)
points(time,n.obs,pch=18,col=2)
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

## analysis
an = ens[2*(1:NT),]
an.stats = apply(an,1,quantile,c(0.025,0.25,0.5,0.75,0.975))
plot(time,an.stats[5,],col="blue",ylim=yrng,xlim=xlim,type='n',ylab="LAI",cex.lab=1.5)
polygon(cbind(c(time, rev(time), 1), c(n.stats[1, ], rev(n.stats[3, ]),
  n.stats[1, 1])), border = NA, col = 'lightblue')
polygon(cbind(c(time, rev(time), 1), c(an.stats[1, ], rev(an.stats[5, ]),
  an.stats[1, 1])), border = NA, col = 'lightgreen')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[2, 1])), border = NA, col = 'green4')
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),  ## Analysis IQR- no disturb
  an0.stats[2, 1])), border = NA, col = 'orange')
lines(time,n.stats[2,],lwd=2,col="blue4")
lines(time,n.true,type='l',lwd=2)
points(time,n.obs,pch=18,col=2)
legend("topleft",legend=c("True","Obs","95%","IQR","Model"),lty=c(1,0,1,1,1),col=c(1,2,"lightgreen","green4","lightblue"),lwd=c(1,NA,15,15,15),cex=1.2,pch=c(NA,18,NA,NA,NA))

```

## Offline State-Space Model

Start by fitting the full state-space model to the data, then move on to figuring out how to approximate that iteratively

```{r}
library(rjags)

SSSD <- "
model{
  q  ~ dgamma(aq,bq)
  r  ~ dgamma(ar,br)
  R  ~ dgamma(aR,bR)
  K  ~ dgamma(aK,bK)
  p  ~ dbeta(ap,bp)
  for(t in 1:nt){
    Y[t] ~ dnorm(n[t],r)
  }
  n[1] <- N0            ## should add some uncertainty
  for(t in 2:nt){
    muN[t] <- n[t-1] + R*n[t-1]*(1-n[t-1]/K)
    muD[t] ~ dnorm(mu0,pa0)
    D[t] ~ dbern(p)
    mu[t] <- D[t]*muD[t] + (1-D[t])*muN[t]
    n[t] ~ dnorm(mu[t],q)
  }
}"
## Run individual models
data = list(Y=n.obs,nt=NT,
            mu0=1,pa0=1,  ## put a bit more confidence in the low value
            aq=0.1,bq=0.1,
            ar=0.1,br=0.1,
            aR=r,bR=1,
            aK=K,bK=1,
            ap=p.dist*50,bp=(1-p.dist)*50, ## try with an informative prior
            N0 = nIC)  
 mod <- jags.model(file=textConnection(SSSD),
                   data=data,
                  n.adapt=1000,n.chains=3,
                  init=list(R = r, K = K, q = n.sd,p=p.dist))
 jdat <- coda.samples(mod,variable.names=c("R","K","p","r","q"),n.iter=3000) ## burnin
 #plot(jdat)
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=30000,thin=10) ## samples
```

```{r}
## summarize output
CI <- apply(as.matrix(jdat),2,quantile,c(0.025,0.5,0.975))
CIN <- CI[,grep(pattern = "n",colnames(CI))]
col=col2rgb("blue")
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim)
ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
lines(CIN[2,],col="blue",lwd=2)

## Disturbance probability
Dsamp <- as.matrix(jdat)
Dsamp <- Dsamp[,grep(pattern = "^D",colnames(Dsamp))]
muD <- apply(Dsamp,2,mean)

CID <- CI[,grep(pattern = "^D",colnames(CI))]
col=col2rgb("blue")
plot(2:NT,muD,ylim = range(CID),xlim=xlim,type='l')
ecoforecastR::ciEnvelope(2:NT,CID[1,],CID[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

if(FALSE){  ## didn't work as expected because D is 0/1, want to plot p[t]
Dden <- data.frame(val=as.vector(Dsamp),year=rep(1:ncol(Dsamp),each=nrow(Dsamp)))
ggplot2::ggplot(Dden, ggplot2::aes(x = val, y = year,  group = year)) +
  ggridges::geom_density_ridges() +
  ggridges::theme_ridges() + 
  ggplot2::theme(legend.position = "none")
}

## Disturb vs Non
CImuN <- CI[,grep(pattern = "muN",colnames(CI))]
CImuD <- CI[,grep(pattern = "muD",colnames(CI))]
colN=col2rgb("green")
colD=col2rgb("orange")
plot(data$Y,ylim = range(rbind(CImuN,CImuD,data$Y[-1])),xlim=xlim,
     ylab="Disturbance Probability")
ecoforecastR::ciEnvelope(2:NT,CImuD[1,],CImuD[3,],col=rgb(colD[1],colD[2],colD[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(2:NT,CImuN[1,],CImuN[3,],col=rgb(colN[1],colN[2],colN[3],0.3*256,maxColorValue=256))

## comparison to EnKF
plot(data$Y,ylim = range(rbind(CIN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[1, ], rev(an0.stats[5, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[1, ], rev(an.stats[5, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
lines(an0.stats[3,],col="orange",lwd=2)
lines(an.stats[3,],col="#b2df8a",lwd=2)
lines(CIN[2,],col="blue",lwd=2)
points(data$Y,pch=18,col=2)
legend("topright",legend=c("EnKF, no disturb","EnKF, disturb","Multinomial","data"),
       lwd=5,col=c("orange","#b2df8a","blue","red"))

```

## Forward Filter version of SSSD, Known Params

```{r}
FilterSSSD <- "
model{
  N ~ dnorm(muIC,paIC)  ## initial conditions
  ## Forecast
  muN <- N + R*N*(1-N/K)
  muD ~ dnorm(mu0,pa0)
  D ~ dbern(p)
  mu <- D*muD + (1-D)*muN
  n ~ dnorm(mu,q)
  ## Analysis
  Y ~ dnorm(n,r)
}"
## Forecast Cycle
CI.fN = matrix(NA,5,NT)
CI.fD = matrix(NA,5,NT)
Xl = nIC         ## set IC
pl = 1/n.sd^2
Dbar = 0
#mu0 = data$mu0   ## set disturbed state
#pa0 = data$pa0
#aqp = data$aq
#bqp = data$bq
#arp = data$ar
#brp = data$br
Analysis <- list()
for(t in 1:NT){
  ## Forecast + Analysis
  update = list(Y=data$Y[t],
                muIC=Xl,paIC=pl,
                mu0=data$mu0,pa0=data$pa0,
                #aq=aqp[t],bq=bqp[t],ar=arp[t],br=brp[t],
                R=r, K=K,p = p.dist, q = 1/n.sd^2, r=1/obs.sd^2     ## hard coding these to start
                )
  mod <- jags.model(file=textConnection(FilterSSSD),
                   data=update,
                  n.adapt=1000,n.chains=3)#,
#                  init=list(X.ic=Xlf))
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000) 
# plot(jdat)

  ## update parameters  
  dat = as.matrix(jdat)
  Xl  = mean(dat[,"n"])
  pl  = 1/var(dat[,"n"])
  CI.fN[,t] = quantile(dat[,"n"],c(0.025,0.25,0.5,0.75,0.975))
  CI.fD[,t] = quantile(dat[,"D"],c(0.025,0.25,0.5,0.75,0.975))
  Dbar[t] = mean(dat[,"D"])
#  mu0[t+1] = mean(dat[,"muD"])
#  pa0[t+1] = 1/var(dat[,"muD"])

#  mq = mean(dat[,"q"])
#  vq = var(dat[,"q"])
#  aqp[t+1] = mq*mq/vq
#  bqp[t+1] = mq/vq
#  mr = mean(dat[,"r"])
#  vr = var(dat[,"r"])
#  arp[t+1] = mr*mr/vr
#  brp[t+1] = mr/vr

  ## save Analysis posterior
  Analysis[[t]] <- dat

}

```

```{r}
plot(0,0,xlim=xlim,ylim=c(0,1))
ecoforecastR::ciEnvelope(time,CI.fD[1,],CI.fD[5,],col=col.alpha(4,0.5))
ecoforecastR::ciEnvelope(time,CI.fD[2,],CI.fD[4,],col=4)
lines(time,CI.fD[2,])
```

### Comparision to EnKF
```{r}
plot(data$Y,ylim = range(rbind(CI.fN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
#ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(1:NT,CI.fN[2,],CI.fN[4,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

lines(an0.stats[3,],col="orange",lwd=2)
lines(an.stats[3,],col="#b2df8a",lwd=2)
#lines(CI.fSSSD[2,],col="blue",lwd=2)
lines(CI.fN[3,],col="blue",lwd=2)
lines(Dbar,lwd=2)
points(data$Y,pch=18,col=2)
legend("topright",legend=c("EnKF, no disturb","EnKF, disturb","Multinomial","data","P(dist)"),
       lwd=5,col=c("orange","#b2df8a","blue","red","black"))
```

### Ridgeline plot of Analysis
```{r}
## organize Analysis state var (n) into a wide format
n.wide <- sapply(Analysis,function(x){x[,"n"]})
colnames(n.wide) <- paste0("t",1:NT)
n.wide <- as.data.frame(n.wide)

## transform to a long format
n.long <- tidyr::pivot_longer(n.wide[,xlim[1]:xlim[2]],
                              cols=1:(diff(xlim)+1),
                              names_to = "time",names_prefix = "t",
                              values_to = "n")
n.long$time <- as.numeric(n.long$time)

## ridge plot
ggplot(n.long) + aes(x=n,y=time,group=time) +
  ggridges::geom_density_ridges(bandwidth = 0.02) +
  labs(x = "LAI", y = "Time")

## violin plot
n.long %>%
 filter(n >= 0 & n <= 8) %>%
 ggplot() +
 aes(x = "", y = n, group = time) +
 geom_violin(adjust = 0.5, scale = "area", fill = "#0c4c8a") +
 labs(x = "Time", y = "LAI") +
 theme_minimal()

```

## Foreward filter, separate process model

In the previous code, both the forecast and analysis steps were done within JAGS. In this version we're pulling out the Forecast step and running that in R since (i) that's how most models work and (ii) it will allow us to more easily distinguish the Forecast prior and Analysis posteriors

```{r}
AnalysisSSSD <- "
model{
  ## prior
  muN ~ dnorm(mufN,pfN)
  muD ~ dnorm(mufD,pfD)
  D ~ dbern(p)
  ## Analysis
  n <- D*muD + (1-D)*muN
  Y ~ dnorm(n,r)
}"
## Storage
CI.fN = matrix(NA,5,NT)
CI.fD = matrix(NA,5,NT)
Dbar = 0
Analysis <- Forecast <- list()
## set IC
nMC = 5000
Xl = nIC         
pl = 1/n.sd^2
Na = rnorm(nMC,Xl,n.sd)
Nf = Na*NA

for(t in 1:NT){
  ## Forecast 
  d.prior <- rbinom(nMC,1,p.dist)   ## prior on disturbance status
  sel <- !as.logical(d.prior)
  Nf[!sel] <- rnorm(sum(d.prior),data$mu0,1/sqrt(data$pa0))  ## draw disturbed predictions
  Nf[sel]  <- rnorm(sum(sel),Na[sel] + r*Na[sel]*(1-Na[sel]/K),n.sd) ## draw undisturbed
  Nf[Nf < 0] <- 0 ## Tobit

  ## priors
  priors <- list(p    = mean(d.prior),
                 mufN = mean(Nf[sel]),
                 mufD = mean(Nf[!sel]),
                 pfN  = 1/var(Nf[sel]),
                 pfD  = 1/var(Nf[!sel])
                 )
  Forecast[[t]] <- c(list(d.prior=d.prior,Nf=Nf),priors)

  ## Analysis
  update = c(list(Y=data$Y[t],r=1/obs.sd^2),priors)
  mod <- jags.model(file=textConnection(AnalysisSSSD),
                   data=update,
                  n.adapt=1000,n.chains=3)
 jdat <- coda.samples(mod,variable.names=c("n","D","muN","muD"),n.iter=10000) 
# plot(jdat)

  ## update parameters  
  dat = as.matrix(jdat)
  Xl  = mean(dat[,"n"])
  pl  = 1/var(dat[,"n"])
  CI.fN[,t] = quantile(dat[,"n"],c(0.025,0.25,0.5,0.75,0.975))
  CI.fD[,t] = quantile(dat[,"D"],c(0.025,0.25,0.5,0.75,0.975))
  Dbar[t] = mean(dat[,"D"])

  ## save Analysis posterior
  Analysis[[t]] <- dat

  ## draw new IC
  Na <- sample(dat[,"n"],nMC,replace=TRUE)

}

```

```{r}
## organize Analysis state var (n) into a wide format
n.wide <- sapply(Analysis,function(x){x[,"n"]})
colnames(n.wide) <- paste0("t",1:NT)
n.wide <- as.data.frame(n.wide)

## transform to a long format
n.long <- tidyr::pivot_longer(n.wide[,xlim[1]:xlim[2]],
                              cols=1:(diff(xlim)+1),
                              names_to = "time",names_prefix = "t",
                              values_to = "n")
n.long$time <- as.numeric(n.long$time)

## ridge plot
ggplot(n.long) + aes(x=n,y=time,group=time) +
  ggridges::geom_density_ridges(bandwidth = 0.02) +
  labs(x = "LAI", y = "Time")

## violin plot
n.long %>%
 filter(n >= 0 & n <= 8) %>%
 ggplot() +
 aes(x = "", y = n, group = time) +
 geom_violin(adjust = 0.5, scale = "area", fill = "#0c4c8a") +
 labs(x = "Time", y = "LAI") +
 theme_minimal()
```

### Comparision to EnKF
```{r}
plot(data$Y,ylim = range(rbind(CI.fN,data$Y)),xlim=xlim,type='n',
     ylab="LAI",xlab="time")
polygon(cbind(c(time, rev(time), 1), c(an0.stats[2, ], rev(an0.stats[4, ]),
  an0.stats[1, 1])), border = NA, col = '#FFBF6C')
polygon(cbind(c(time, rev(time), 1), c(an.stats[2, ], rev(an.stats[4, ]),
  an.stats[1, 1])), border = NA, col = ecoforecastR::col.alpha('#b2df8a',0.7))
#ecoforecastR::ciEnvelope(1:NT,CIN[1,],CIN[3,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))
ecoforecastR::ciEnvelope(1:NT,CI.fN[2,],CI.fN[4,],col=rgb(col[1],col[2],col[3],0.3*256,maxColorValue=256))

lines(an0.stats[3,],col="orange",lwd=2)
lines(an.stats[3,],col="#b2df8a",lwd=2)
#lines(CI.fSSSD[2,],col="blue",lwd=2)
lines(CI.fN[3,],col="blue",lwd=2)
lines(Dbar,lwd=2)
points(data$Y,pch=18,col=2)
legend("topright",legend=c("EnKF, no disturb","EnKF, disturb","Multinomial","data","P(dist)"),
       lwd=5,col=c("orange","#b2df8a","blue","red","black"))

### Save outputs required to include figure in Results
save(data,CI.fN,time,an0.stats,an.stats,NT,Dbar,xlim,col,file="SSSD.RData")

```

## Conceptual Figure
```{r, echo=FALSE}
## params
set.seed(1066)
p.dist = 0.05
n0 = 1
nIC = 6
r = .1
K = 6
n.sd = 0.25
NE = 1000
obs=1
obs.sd = 0.75

## generate IC ensemble
ens = rnorm(NE,nIC,n.sd)

## Forecast step
LAI = ens + r*ens*(1-ens/K) + rnorm(NE,0,n.sd) 
disturbed = as.logical(rbinom(NE,1,p.dist))
LAI[disturbed] = abs(rnorm(sum(disturbed),n0,0.5)) ## disturbance

##Approximate forecast with Normal
xseq = seq(0,8,length=1000)
muf = mean(LAI)
Pf = var(LAI)

## analysis step: combine previous forecast with observed data
H <- matrix(1,ncol=1)
K <- Pf * t(H) %*% solve(H%*%Pf%*%t(H) + obs.sd^2)
mua <- muf + K*(obs - muf) 
Pa <- (1-K %*% H)*Pf

## EnKF
hist(LAI,probability = TRUE)#,ylim=c(0,dnorm(mua,mua,sqrt(Pa))))
lines(xseq,dnorm(xseq,muf,sqrt(Pf)),col="#b2df8a",lwd=3)
lines(xseq,dnorm(xseq,obs,obs.sd),col='red',lwd=3)
lines(xseq,dnorm(xseq,mua,sqrt(Pa)),col='orange',lwd=3)
legend("topleft",legend=c("forecast","data","assimilation"),lwd=3,col=c("#b2df8a","red","orange"))

## Forecast: Mixture of normals
mu1 = mean(LAI[!disturbed])
P1 = var(LAI[!disturbed])
mu2 = mean(LAI[disturbed])
P2 = var(LAI[disturbed])

## assimilation
priors <- list(p = p.dist,
                 mufN = mu1,
                 mufD = mu2,
                 pfN  = 1/P1,
                 pfD  = 1/P2,
                 Y = obs,
                 r = 1/obs.sd^2
                 )
mod <- jags.model(file=textConnection(AnalysisSSSD),
                   data=priors,n.adapt=1000,n.chains=3)
jdat <- coda.samples(mod,variable.names=c("n"),n.iter=10000)

hist(LAI,probability = TRUE,ylim=c(0,dnorm(mu1,mu1,sqrt(P1))))
lines(xseq,(1-p.dist)*dnorm(xseq,mu1,sqrt(P1)),col="#b2df8a",lwd=3)
lines(xseq,p.dist*dnorm(xseq,mu2,sqrt(P2)),col="brown",lwd=3)
lines(xseq,dnorm(xseq,obs,obs.sd),col='red',lwd=3)
lines(density(as.matrix(jdat)),col="blue",lwd=3)
legend("top",legend=c("undisturbed","disturbed","data","assimilation"),lwd=3,col=c("#b2df8a","brown","red","blue"))

#legend("top",legend=c("undisturbed","disturbed"),lwd=3,col=c("#b2df8a","brown"))

```

## Todo
-   Tweak model to have a simple C budget -- where does the C go when the model doesn't assimilate disturbance?
-   Calibrate state and transition probabilities for a single region
-   Upgrade code to multinomial
-   Add PFT switching
-   Apply to SIPNET SDA
    - version of code into PEcAn
    - test coupling to SIPNET
-   Real-world data examples
-   Add ensemble weighting so that we can manage ensemble sizes
-   Add "fusion" if a case becomes too rare to worry about
-   Think about whether we need to add LULC state to state variables & how LULC state would be assimilated
-   Test against subseasonal dynamics (e.g. distinguishing phenology from disturbance)
-   investigate whether we need a "jumpback" capacity once disturbance event type and timing is resolved
-