### 2023-01-12 discussion.Rmd
knitr debugging

# Discussion

* Interpretation
  * TBD based on results

* Issues
  * Need a more detail exploration of rate of false pos/neg: affected by p.dist, R, Q, P0 -- how much do we need to simulate?
  * Would like to add look-back capabilities (ensemble backward smoothing) to update assignments based on more recent info
    * related: with multi-PFT need to check for, and protect against, long term PFT leakage
  * Computational demand it high for problems with many grid cells
    * Efficient (semi-)analytic approximations or LUT
    * work at larger spatial scales and account for subgrid disturbance heterogeneity implicitly [@McCabe2019]
  * Multisite needs a way of estimating $\rho$ by site with cross site correlations, rather than unique $\rho$ for all permutations
    * for $\rho$ handle the covariances of nearby sites -- disturbance in one usually increases the probability of disturbance in others
    * similarly, need to be able to compose Multisite $\mu_f$ and $P_f$, from individual site values with reasonable approximations of cross-[site x pool] correlations

* Discussion of similar approaches
[@Ye2020]

* Calibration was "quick and dirty" because the focus was on disturbance detection not default dynamics. For actual carbon monitoring and forecasting we would recommend a more extensive multi-site Bayesian calibration so that parameter uncertainty can be propagated [CITE istem 2018, dietze 2017], preferably employing hierarchical approaches to also capture parameter heterogeneity [CITE istem 2021]. Similarly, would recommend a more robust estimation and propagation of process error as part of the data assimilation approach itself [CITE raiho].

* Implications  (borrow text from CMS proposals)
  * Improving MRV
  * Recovery forecasting
  * Disturbance nowcasting
    * requires reducing temporal resolution

* Future directions
  * Disturbance priors need to be spatially heterogenous and evolve temporally
    * Connect to state-and-transition literature
    * shared spatial information (e.g. dispersal kernels)
  * Scaling up -> spatially implicit
    * ED2
    * Contagion

* OREGON: FUTURE DIR
  * disturbance products are 'seeing' more than is in AGB product alone
    * also would benifit from more products (e.g. LAI)
  * would be good to add 'memory' to algorithm so further evidence of disturbance reinforces initial detection (similar to BLC); would allow reduction in time step and greater sensor fusion

* Run just the Assimilation step for simulated disturbance that vary in magnitude, process error, observation error, etc to detect sensitivity
* Approximation of not simulating disturbed ensemble members and relying SOLEY on ensemble adjustment
* Analytical approximation of sequential beta-binomial then KF vs full joint model

Leftover text from Methods:
  + x SS Multi-Disturbance - distinct
  + x SS Multi-Disturbance w/ distinct, PFT switching
      * mention parallel PFT ensembles
      * need to protect against long-term PFT leakage; related to more general need to "look back"
  + x Simple Multi-site: spatial cov
  + Future
      * SSMD - initially overlapping but diverge
        * how to handle updating of assignment based on new information?
      * SSMD - transition probabilities change with state
        * rate of false pos/neg: affected by p.dist, R, Q, P0 -- how much do we need to simulate?